### 2023-11-13 article/article.tex

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
\PassOptionsToPackage{dvipsnames,svgnames,x11names}{xcolor}
%
\documentclass[
  authoryear,
  review,
  1p]{elsarticle}

\usepackage{amsmath,amssymb}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
\usepackage{lmodern}
\ifPDFTeX\else  
    % xetex/luatex font selection
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\setlength{\emergencystretch}{3em} % prevent overfull lines
\setcounter{secnumdepth}{5}
% Make \paragraph and \subparagraph free-standing
\ifx\paragraph\undefined\else
  \let\oldparagraph\paragraph
  \renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
  \let\oldsubparagraph\subparagraph
  \renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother

\tnotetext[t1]{The opinions expressed in this paper are those of the authors and do not necessarily reflect the views of the World Bank, its Board of Directors, or the countries it represents.}
\makeatletter
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{Figure}
\else
  \newcommand\figurename{Figure}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother
\makeatletter
\@ifpackageloaded{tcolorbox}{}{\usepackage[skins,breakable]{tcolorbox}}
\makeatother
\makeatletter
\@ifundefined{shadecolor}{\definecolor{shadecolor}{rgb}{.97, .97, .97}}
\makeatother
\makeatletter
\makeatother
\makeatletter
\makeatother
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\usepackage[]{natbib}
\bibliographystyle{elsarticle-harv}
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={Quantile Regressions via Method of Moments with multiple fixed effects },
  pdfauthor={Fernando Rios-Avila; Leonardo Siles; Gustavo Canavire-Bacarreza},
  pdfkeywords={Fixed effects, Linear heteroskedasticity, Location-scale
model, Quantile regression},
  colorlinks=true,
  linkcolor={blue},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue},
  pdfcreator={LaTeX via pandoc}}

\setlength{\parindent}{6pt}
\begin{document}

\begin{frontmatter}
\title{Quantile Regressions via Method of Moments with multiple fixed
effects \tnoteref{t1}}
\author[1]{Fernando Rios-Avila%
\corref{cor1}%
}
 \ead{friosavi@levy.org} 
\author[2]{Leonardo Siles%
%
}
 \ead{leonardosiles@hotmail.com} 
\author[3]{Gustavo Canavire-Bacarreza%
%
}
 \ead{gcanavire@worldbank.org}

\affiliation[1]{organization={Levy Economics Institute of Bard
College},,postcodesep={}}
\affiliation[2]{organization={Universidad Católica
Boliviana},,postcodesep={}}
\affiliation[3]{organization={World Bank},,postcodesep={}}

\cortext[cor1]{Corresponding author}

\begin{abstract}
This paper proposes a new method to estimate quantile regressions with
multiple fixed effects. The method, which expands on the strategy
proposed by \citet{mss2019}, allows for the inclusion of multiple fixed
effects and provides various alternatives for the estimation of standard
errors. We provide Monte Carlo simulations to show the finite sample
properties of the proposed method in the presence of two sets of fixed
effects. Finally, we apply the proposed method to estimate the
determinants of the surplus of government as a share of GDP, allowing
for both time and country fixed effects.
\end{abstract}

\begin{keyword}
    Fixed effects \sep Linear heteroskedasticity \sep Location-scale
model \sep 
    Quantile regression
\end{keyword}
\end{frontmatter}
    \ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[borderline west={3pt}{0pt}{shadecolor}, breakable, frame hidden, interior hidden, enhanced, boxrule=0pt, sharp corners]}{\end{tcolorbox}}\fi

\hypertarget{introduction}{%
\section{Introduction}\label{introduction}}

Quantile regression (QR), introduced by \citet{koenkerbasset1978}, is an
estimation strategy used for modeling the relationships between
explanatory variables X and the conditional quantiles of the dependent
variable \(Q_y (\tau|x)\). Using QR one can obtain richer
characterizations of the relationships between dependent and independent
variables, by accounting for otherwise unobserved heterogeneity.

A relatively recent development in the literature has focused on
extending quantile regressions analysis to include individual fixed
effects in the framework of panel data. However, as described in
\citet{neymanscott1948}, and \citet{lancaster2000}, when individual
fixed effects are included in quantile regression analysis an incidental
parameter problem is generated. While many strategies have been proposed
for estimating this type of model (see \citet{galvao2017quantile} for a
brief review), neither has become standard because of their restrictive
assumptions in regard to the inclusion of individual and multiple fixed
effects, the computational complexity, and implementation.

More recently, \citet{mss2019} (MSS hereafter) proposed a methodology
based on a conditional location-scale model, similar to the one
described in \citet{he1997} and \citet{zhao2000}, for the estimation of
quantile regressions models for panel data via a method of moments. This
method allows individual fixed effects to have heterogeneous effects on
the entire conditional distribution of the outcome, rather than
constraining their effect to be a location shift only, as in
\citet{canay2011}, \citet{koenker2004}, and \citet{lamarche2010}.

In principle, under the assumption that data-generating process behind
the data is based on a multiplicative heteroskedastic process that is
linear in parameters \citep[\citet{he1997}, \citet{zhao2000},
\citet{cameron2005}]{mss2019}, the effect of a variable \(X\) on the
\(q_{th}\) quantile can be derived as the combination of a location
effect, and scale effect moderated by the quantile of an underlying
i.i.d. error. For statistical inference, MSS derives the asymptotic
distribution of the estimator, suggesting the use of bootstrap standard
errors, as well.

This methodology is not meant to substitute for the use of standard
quantile regression analysis. That said, given the assumptions required
for the identification of the model, it provides a simple and fast
alternative for the estimation of quantile regression models with
individual fixed effects.

In this framework, our paper expands on \citet{mss2019}, following some
of the suggestions by these authors regarding further research. First,
making use of the properties of generalized method of moments (GMM)
estimators, we derive various alternatives for the estimation of
standard errors based on the empirical influence functions of the
estimators. Second, we reconsider the application of Frisch-Waugh-Lovell
(FWL) theorem \citep[ and \citet{lovell1963}]{frishwaugh1933} to extend
the MSS estimator to allow for the inclusion of multiple fixed effects,
for example, individual and year fixed effects.

The rest of the paper is structured as follows: section 2 presents the
basic setup of the location-scale model described in \citet{he1997} and
\citet{zhao2000}, tying the relationship between the standard quantile
regression model and the location-scale model. It also revisits the
methodology of MSS, proposing alternative estimators for the standard
errors based on the properties of GMM estimators and the empirical
influence functions. It also shows that the FWL theorem can be used to
control for multiple fixed effects. Section 3 presents the results of a
small simulation study and section 4 illustrates the application of the
proposed methods with two empirical examples. Section 5 concludes.

\hypertarget{sec-method}{%
\section{Methodology}\label{sec-method}}

\hypertarget{sec-betas}{%
\subsection{Quantile Regression: Location-Scale model}\label{sec-betas}}

Quantile regressions are used to identify relationships between the
explanatory variables \(X\) and the conditional quantiles of the
dependent variable \(Q_y(\tau|X)\). This relationship is commonly
assumed to follow a linear functional form:

\begin{equation}\protect\hypertarget{eq-eq1}{}{Q_y(\tau|X) =X\beta(\tau)
}\label{eq-eq1}\end{equation}

This allows for nonlinearities in the effect of \(X\) on \(Y\) across
all values of \(\tau\). This formulation can also be related to a random
coefficient model, in which all coefficients are assumed to be some
nonlinear function of \(\tau\) where \(\tau\) follows a random uniform
distribution.

An alternative formulation of quantile regressions is the location-scale
model. This approach assumes that the conditional quantile of \(Y\)
given \(X\) and \(\tau\) can be expressed as a combination of two
models: the location model, which describes the central tendency of the
conditional distribution; and the scale model, which describes
deviations from the central tendency:

\begin{equation}\protect\hypertarget{eq-eq2}{}{Q_y(\tau|X) =X\beta+X\gamma(\tau)
}\label{eq-eq2}\end{equation}

Here, the location parameters \(\beta\) are typically identified using a
linear regression model (as in \citet{mss2019}) or a median regression
(as in \citet{zhao2000}) and the scale parameters \(\gamma(\tau)\) can
be estimated using standard approaches.

Both the standard quantile regression (Equation~\ref{eq-eq1}) and the
location-scale specification (Equation~\ref{eq-eq2}) can be estimated as
the solution to a weighted minimization problem:

\begin{equation}\protect\hypertarget{eq-eq3}{}{\hat{\beta}(\tau) = \underset{\beta}{\operatorname{argmin}}
\left( \sum_{i\in y_i\geq x_i'\beta} \tau (y_i - x_i'\beta) - \sum_{i\in y_i<x_i'\beta} (1-\tau)(y_i - x_i'\beta) \right)
}\label{eq-eq3}\end{equation}

One characteristic of this estimator is that the \(\beta(\tau)\)
coefficients are identified locally and thus the estimated quantile
coefficients will exhibit considerable variation when analyzed across
\(\tau\). It is also implicit that if one requires an analysis of the
entire distribution, it would be necessary to estimate the model for
each quantile.\footnote{There are other estimators that provide smoother
  estimates for the quantile regression coefficients using a kernel
  local weighted approach \citep{kaplan2017}, as well as identifying the
  full set of quantile coefficients while simultaneously assuming some
  parametric functional forms \citep{frumentobotai2016}.}

One insightful extension to the location-scale parameterizations
suggested by \citet{he1997}, \citet{cameron2005}, and \citet{mss2019} is
to assume that the data-generating process (DGP) can be written as a
linear model with a multiplicative heteroskedastic process that is
linear in parameters.\footnote{\citet{mss2019} also discuss a model
  where heteroskedasticity can be an arbitrary nonlinear function
  \(\sigma(x_i'\gamma)\), but develop the estimator for the linear case,
  i.e., when \(\sigma()\) is the identity function.}

\begin{equation}\protect\hypertarget{eq-eq4}{}{\begin{aligned}
y_i &=x_i'\beta+\nu_i \\
\nu_i &=\varepsilon_i \times x_i'\gamma 
\end{aligned}
}\label{eq-eq4}\end{equation}

Under the assumption that \(\varepsilon\) is an i.i.d. unobserved random
variable that is independent of \(X\), the conditional quantile of \(Y\)
given \(X\) and \(\tau\) can be written as

\begin{equation}\protect\hypertarget{eq-eq5}{}{Q_y(\tau|X) =X\beta+Q_\varepsilon(\tau) \times X\gamma 
}\label{eq-eq5}\end{equation}

In this setup, the traditional quantile coefficients are identified as
the location model coefficients plus the scale model coefficients
moderated by the \(\tau_{th}\) unconditional quantile of the
standardized error \(\varepsilon\). For simplicity we will use
\(q(\tau)\) to denote \(Q_\varepsilon(\tau)\) in the rest of the paper.

\begin{equation}\protect\hypertarget{eq-eq6}{}{\beta(\tau) = \beta + q(\tau) \times \gamma 
}\label{eq-eq6}\end{equation}

While this specification imposes a strong assumption on the DGP, it has
two advantages over the standard quantile regression model. First,
because the location-scale model can be identified globally, with only a
single parameter (\(q(\tau)\)) requiring local estimation, this
estimation approach will be more efficient than the standard quantile
regression model (\citet{zhao2000}). Second, under the assumption that
\(X\gamma\) is strictly positive, the model will produce quantile
coefficients that do not cross.

Following MSS, the quantile regression model defined by
Equation~\ref{eq-eq5} can be estimated using a generalized method of
moments approach. And while it is possible to identify all coefficients
(\(\beta,\gamma, q(\tau)\)) simultaneously, we describe and use the
implementation approach advocated by MSS, which identifies each set of
coefficients separately.

First, the location model can be estimated using a standard linear
regression model, where the dependent variable is the outcome \(Y\) and
the independent variables are the explanatory variables \(X\) (including
a constant) with an error \(u\), which is by definition heteroskedastic.
In this case, the location-model coefficients are identified under the
following condition:

\begin{equation}\protect\hypertarget{eq-eq7}{}{\begin{aligned}
      y_i &=x_i'\beta+\nu_i \\
      E\big[ x_i \nu_i\big] &=0
      \end{aligned}
}\label{eq-eq7}\end{equation}

Second, after the location model is estimated, the scale coefficients
can be identified by modeling heteroskedasticity as a linear function of
characteristics \(X\). For this we use the absolute value of the errors
from the location model \(u\) as dependent variable, which allows us to
estimate the conditional standard deviation (rather than conditional
variance) of the errors. In this case, the coefficients are identified
under the following condition:

\begin{equation}\protect\hypertarget{eq-eq8}{}{\begin{aligned}
  |\nu_i| &=x_i'\gamma+\omega_i \\
   E\big[ x_i (|\nu_i| -x_i'\gamma ) \big] &=0
  \end{aligned}
}\label{eq-eq8}\end{equation}

Third, given the location and scale coefficients, the \(\tau_{th}\)
quantile of the error \(\varepsilon\) can be estimated using the
following condition:

\begin{equation}\protect\hypertarget{eq-eq9}{}{\begin{aligned}
  E\left[  \mathbb{1}\left(x_i' (\beta +   \gamma q(\tau)) \geq y_i \right) - \tau \right] &=0  \\
  E\left[  \mathbb{1}\left(   q(\tau) \geq \frac{y_i-x_i'\beta}{x_i'\gamma} \right) - \tau \right] &=0  \\
  \end{aligned}
}\label{eq-eq9}\end{equation}

where one identifies the quantile of the error \(\varepsilon\) using
standardized errors \(\frac{y_i-x_i'\beta}{x_i'\gamma}\) or by finding
the values that identify the overall quantile coefficients
\(\beta(\tau)=\beta + \gamma q(\tau)\). Afterwards, the conditional
quantile coefficients are simply defined as the combination of the
location and scale coefficients.

\hypertarget{sec-se}{%
\subsection{Standard Errors: GLS, Robust, Clustered}\label{sec-se}}

As discussed in the previous section, the estimation of quantile
regression coefficients using the location-scale model with
heteroskedastic linear errors can be estimated using a the following set
of moments, which fits within the GMM framework:

\begin{equation}\protect\hypertarget{eq-eq10}{}{\begin{aligned}
  E[x_i \nu_i  ] &= E[h_{1,i}]=0 \\
  E[x_i  (|\nu_i|-x_i \gamma) ] &=E[h_{2,i}]=0 \\
  E\left[  \mathbb{1}\left(   q(\tau)\geq \frac{y_i-x_i'\beta}{x_i'\gamma} \right) - \tau \right] 
  &=E[h_{3,i}]=0 
  \end{aligned}
}\label{eq-eq10}\end{equation}

Under the conditions described in \citet{newey_chapter_1994} (see
section 7), \citet{cameron2005} (see chapter 6.3.9), or as shown in
\citet{mss2019}, the location, scale, and residual quantile coefficients
are asymptotically normal.\footnote{\citet{zhao2000} also shows that the
  quantile coefficients for the location-scale model follow a normal
  distribution, but uses the assumption that the location model is
  derived using a least absolute deviation approach (median regression).}

Call \(\theta=[ \beta' \ \ \gamma' \ \ q(\tau) ]'\) the set of
coefficients that are identified by the moment conditions in
Equation~\ref{eq-eq10}, a just identified model, and the function
\(h_i\) is a vector function that stacks all the moments described in
Equation~\ref{eq-eq10} at the individual level. Then \(\hat\theta\)
follows a normal distribution with mean \(\theta\) and
variance-covariance matrix \(V(\theta)\) that is estimated as

\[
\hat{V}(\hat\theta)=\frac{1}{N} 
\bar G(\hat\theta)^{-1} 
\left( \frac{1}{N} \sum_{i=1}^N h_i h_i'  \Big|_{\theta=\hat\theta} \right) 
\bar G(\hat\theta)^{-1} 
\]

which is equivalent to the Eicker-White heteroskedasticity-consistent
estimator for least-squares estimators.

Here, the inner product is the moment covariance matrix and
\(\bar{G}(\theta)\) is the Jacobian matrix of the moment equations
evaluated at \(\hat\theta\).

\[\bar{G}(\theta) =-\frac{1}{N} \sum_{i=1} \frac{\partial h_i}{\partial \theta'} \Big|_{\theta=\hat\theta}\]

In this framework, the quantile regression coefficients, a combination
of the location-scale-quantile estimates, will follow a normal
distribution with mean
\(\beta(\tau) = \beta+\gamma q(\varepsilon | \tau)\) and
variance-covariance matrix equal to

\[\hat{V}(\beta(\tau)) = \Xi \hat{V}(\hat\theta) \Xi'
\]

where \(\Xi\) is a \(k \times (2k+1)\) matrix defined as

\begin{equation}\protect\hypertarget{eq-eqqtile}{}{\Xi = [ I(k), \hat q(\tau) \times I(k), \hat \gamma ]
}\label{eq-eqqtile}\end{equation}

with \(I(k)\) being an identity matrix of dimension \(k\) (number of
explanatory variables in \(X\) including the constant).

While it is possible to estimate the variance-covariance matrix using
simultaneous model estimation for a just identified model, it is more
efficient to estimate each set of coefficients separately. Afterward,
the variance-covariance matrix can be estimated using the empirical
influence functions of the estimators (see \citet{jann_2020} for an
overview of the application and \citet{hampel2005} for an in-depth
review).

Specifically, given an arbitrary vector of empirical influence functions
\(\lambda_i(\theta)\), the variance-covariance matrix can be estimated
as

\begin{equation}\protect\hypertarget{eq-eqvcv}{}{\hat{V}(\theta) = \frac{1}{N^2} \sum_{i=1}^N \lambda_i(\theta) \lambda_i(\theta)'
}\label{eq-eqvcv}\end{equation}

where the influence functions are defined as:

\[\lambda_i(\theta) = \bar G(\theta)^{-1} h_i(\theta)\]

For the specific case of quantile regressions through moments, the
influence functions for the location, scale, and quantile coefficients
are\footnote{The derivation of the influence functions can be found in
  the appendix.}

\begin{equation}\protect\hypertarget{eq-ifs}{}{\begin{aligned}
\lambda_i(\theta)&=
  \begin{bmatrix}
  \lambda_{i}(\beta) \\
  \lambda_{i}(\gamma) \\
  \lambda_{i}(q(\tau)) 
  \end{bmatrix} \\
\lambda_{i}(\beta)&=N (X'X)^{-1}  x_i ( x_i'\gamma) \times \varepsilon_i \\
\lambda_{i}(\gamma)&= N(X'X)^{-1} x_i ( x_i' \gamma ) \times (\tilde \varepsilon_i -1\big)\\
\lambda_{i}(q(\tau))&=\frac{\tau-\mathbb{1}\big( q(\tau)  \geq \varepsilon_i  \big) }{f_{\varepsilon}(q(\tau))}
- \frac{ x_i'\gamma \times \varepsilon_i }{\bar x_i'\gamma} 
-  q(\tau) \frac{  x_i' \gamma  \times (\tilde \varepsilon_i -1\big)}{\bar x_i'\gamma}
\end{aligned}
}\label{eq-ifs}\end{equation}

The different types of standard errors estimation thus depend on the
assumptions imposed for the estimation of \(V(\theta)\).

\hypertarget{robust-standard-errors}{%
\subsubsection{Robust Standard Errors}\label{robust-standard-errors}}

The first and most natural standard error estimator is given by equation
Equation~\ref{eq-eqvcv}. This is equivalent to the Eicker-White
heteroskedasticity-consistent estimator for least-squares estimators.
Considering the location-scale model, the variance-covariance matrix for
the quantile coefficients can be estimated as

\[\hat{V}_{robust}
  \begin{pmatrix}
  \hat\beta \\
  \hat\gamma \\
  \hat q(\tau)
  \end{pmatrix}
 = \frac{1}{N^2} 
 \begin{pmatrix}
 \sum \lambda_i(\beta) \lambda_i(\beta)' & \sum \lambda_i(\beta) \lambda_i(\gamma)' & \sum \lambda_i(\beta) \lambda_i(q(\tau))' \\
 \sum \lambda_i(\gamma) \lambda_i(\beta)' & \sum \lambda_i(\gamma) \lambda_i(\gamma)' & \sum \lambda_i(\gamma) \lambda_i(q(\tau))' \\
 \sum \lambda_i(q(\tau)) \lambda_i(\beta)' & \sum \lambda_i(q(\tau))\lambda_i(\gamma)' & \sum \lambda_i(q(\tau)) \lambda_i(q(\tau))' 
  \end{pmatrix}
\]

This estimator of standard errors should be robust to arbitrary
heteroskedasticity. However, because the location-scale specification
relies on the correct specification of the model heteroskedasticity,
large differences in standard errors compared to GLS-standard errors may
be an indication of misspecification of the model.

\hypertarget{clustered-standard-errors}{%
\subsubsection{Clustered Standard
Errors}\label{clustered-standard-errors}}

Because one of the typical applications of quantile regressions is the
analysis of panel data, allowing for clustered standard errors at the
individual level is important. If the unobserved error \(\varepsilon\)
is correlated within clusters, generalized least squares (GLS) standard
errors could be severely biased. The standard recommendation has been to
report block-bootstrap standard errors, clustered at the individual
level.

Because we have access to the influence functions, it is straightforward
to estimate one-way clustered standard errors.

Let \(N_G\) be the total number of clusters \(g\), where
\(g=1\dots N_G\). The clustered variance covariance matrix is given
by\footnote{It should be noted that one could just as well apply the
  insights of \citet{cameron_robust_2011}, allowing for multiway
  clustering.}

\[\hat{V}_{clustered}
  \begin{pmatrix}
  \hat\beta \\
  \hat\gamma \\
  \hat q(\tau)
  \end{pmatrix}
 = \frac{1}{N^2} 
 \begin{pmatrix}
 \sum_{g=1}^{N_G} S\lambda_i(\theta) S\lambda_i(\theta)'
  \end{pmatrix}
\]

where \(S\lambda_i(\theta)\) is the sum of the influence functions over
all observations within a given cluster \(g\).

\[S\lambda_i(\theta) = \sum_{i\in g} \lambda_i(\theta)
\]

\hypertarget{gls-standard-errors}{%
\subsubsection{GLS Standard Errors}\label{gls-standard-errors}}

The standard errors proposed by MSS can be understood as an application
of GLS, which will be valid as long as the model for heteroskedasticity
is correctly specified.\footnote{As discussed in most econometric
  textbooks, for example \citet{cameron2005}, one approach to correct
  for heteroskedasticity, when the heteroskedasticity functional form is
  known or can be estimated, is to use weighted least squares. While
  feasible this approach would defeat the purpose of identifying
  quantile effects exploiting the heteroskedasticity of the model.} To
estimate the GLS standard errors, we make use of the following property:

Consider the influence functions and robust variance-covariance matrix
for the location coefficients:

\[\begin{aligned}
\hat V(\hat \beta) &= \frac{1}{N} \sum_i^N \lambda_{i}(\beta) \lambda_{i}(\beta)' \\
&= \frac{1}{N}  (X'X)^{-1}  \sum_i^N x_i x_i' ( x_i'\gamma \times \varepsilon_i)^2 (X'X)^{-1} \\
\end{aligned}
\]

Under the assumption that the model for heteroskedasticity is correctly
specified, we can apply the law of iterated expectations and rewrite the
variance-covariance matrix as

\[\begin{aligned}
\hat V(\hat \beta) &= \frac{1}{N} \sum_i^N \lambda_{i}(\beta) \lambda_{i}(\beta)' \\
&= E(\varepsilon_i^2) \frac{1}{N}  (X'X)^{-1}  \sum_i^N x_i x_i' ( x_i'\gamma)^2 (X'X)^{-1} \\
&= \sigma^2_{\varepsilon} \frac{1}{N}  (X'X)^{-1}  \hat\Omega_{\beta\beta} (X'X)^{-1}  
\end{aligned}
\]

This standard error estimator is an application of GLS that accounts for
the heteroskedasticity the model uses to identify the quantile
coefficients. We can apply the same principle to find the GLS standard
errors for the system of location-scale and quantile coefficients. To do
this, we define the following modified influence functions:

\[\begin{aligned}
\tilde \lambda_{1,i}&=\tilde \lambda_{2,i}=N (X'X)^{-1}  x_i ( x_i'\gamma) \\  
\tilde \lambda_{3,i}&=x_i' \gamma  \\
 \psi_{i,1} &= \varepsilon_i \\
 \psi_{i,2} &= \tilde \varepsilon_i -1 \\
 \psi_{i,3} &= \frac{1}{ x_i' \gamma }\frac{\tau-\mathbb{1}\big( q(\tau)  \geq \varepsilon_i  \big) }{ f_{\varepsilon}(q(\tau))} - \frac{\varepsilon_i }{\bar x_i'\gamma} 
-  q(\tau) \frac{(\tilde \varepsilon_i -1\big)}{\bar x_i'\gamma}
\end{aligned}
\]

Then, the GLS standard errors for the location-scale and quantile
coefficients can be estimated as

\[\hat{V}_{gls}
  \begin{pmatrix}
  \hat\beta \\
  \hat\gamma \\
  \hat q(\tau)
  \end{pmatrix}
 = \frac{1}{N^2} 
 \begin{pmatrix}
 \hat\sigma_{11} \hat\Omega_{11} & \hat\sigma_{12} \hat\Omega_{12} & \hat\sigma_{11} \hat\Omega_{13} \\
 \hat\sigma_{21} \hat\Omega_{21} & \hat\sigma_{22} \hat\Omega_{22} & \hat\sigma_{21} \hat\Omega_{23} \\
 \hat\sigma_{31} \hat\Omega_{31} & \hat\sigma_{32} \hat\Omega_{32} & \hat\sigma_{33} \hat\Omega_{33}
  \end{pmatrix}
\]

where

\[\begin{aligned}
\hat\Omega_{jk} &= \frac{1}{N} \sum_{i=1}^N \tilde \lambda_{i,j}(\theta) \tilde \lambda_{i,k}(\theta)' \\
\hat\sigma_{jk} &= \frac{1}{N} \sum_{i=1}^N \psi_{i,j} \psi_{i,k}
\end{aligned}
\]

This estimator of stand ard errors is equivalent to the one derived by
MSS using Theorem 3.

\hypertarget{multiple-fixed-effects-expanding-on-mss2019}{%
\subsection{\texorpdfstring{Multiple Fixed Effects: Expanding on
\citet{mss2019}}{Multiple Fixed Effects: Expanding on @mss2019}}\label{multiple-fixed-effects-expanding-on-mss2019}}

Using the setup described in the previous section, \citet{mss2019}
proposes an extension to the model proposed by \citet{he1997} that
enables the estimation of quantile regression models with panel data,
allowing for the inclusion of individual fixed effects. However, as the
authors suggest, the methodology can be generalized to allow for the
inclusion of multiple fixed effects. This type of analysis can be useful
when considering data such as employer-employee linked data
\citep{abowed2006} or teacher-student linked data
\citep{harrissass2011}. Or, in the most common case, allowing to control
for both individual and time fixed effects.

We return to the original model and now assume there are sets of
unobserved heterogeneity that are assumed to be constant across
observations, if they belong to common groups. In panel data, the groups
would be the individual fixed effects and the time fixed effects.
Without loss of generality, we can assume that the data-generating
process is as follows:

\[\begin{aligned}
  y_{i} &= x_{i}' \beta + \delta_{g1} + \delta_{g2} + \nu_i \\
  \nu_i &= \varepsilon_i \times (x_{i}' \gamma + \zeta_{g1} + \zeta_{g2})   
  \end{aligned}
\]

where we assume \(x_{i}\) vary across groups \(g_1\) and \(g_2\) (and
thus are not collinear) and that \(\delta's\) and \(\zeta's\) are the
location and scale effects associated with groups fixed
effects.\footnote{We could just as well consider multiple sets of fixed
  effects.}

If the dimension of groups \(g_k\) is low, this model could be estimated
using a dummy inclusion approach following Section~\ref{sec-betas}, and
the standard errors obtained as discussed in Section~\ref{sec-se}.
However, if the dimensionality of \(g_k\) is high, the dummy inclusion
approach may not be feasible. A more feasible approach is to apply the
FWL theorem and partial out the impact of the group fixed effects on the
control variables \(x_{i}\), and the outcome of interest \(y_{i}\),
using a similar approach for the identification of \(\sigma(x)\). In the
case of unbalanced setups with multiple groups, the estimation involves
iterative processes for which various approaches have been suggested and
implemented (see for example, \citet{correia_feasible_nodate},
\citet{gaure2013}, \citet{rios2015}, among others).

When applying the partialing-out approach, some modifications to the
approach described in Section~\ref{sec-betas} are needed.

First, for all dependent and independent variables in the model
(\(w=y,x\)), we partial out the group fixed effects and obtain the
centered-residualized variables:

\[\begin{aligned}
w_{i} &= \delta_{g1}^w + \delta_{g2}^w + u_{i}^w \\
w_{i}^{rc} &= E(w_{i}) + \hat{u}_{i}^w
\end{aligned}
\]

Afterward, we estimate the location model using the
centered-residualized variables:\footnote{Using centered-residualized
  variables allows us to include a constant in the model specification,
  which simplifies the derivation of the influence functions. However,
  as with other fixed effects models, the constant is not identified and
  thus should not be interpreted.}

\[y_{i}^{rc} = x_{i}^{rc'} \beta + \nu_{i}
\]

Because \(|\hat \nu_i|\) is the dependent variable for the scale model,
we apply the partialing out and recentering to this expression
(\(|\hat \nu_i|^{rc}\)), and use that to estimate the following model:

\[|\hat\nu_{i}|^{rc} = x_{i}^{rc'} \gamma + \omega_{i}\]

Finally, the standardized residuals \(\varepsilon_i\) can be obtained as
follows:

\[\hat{\varepsilon}_{i} = \frac{\nu_{i}}{|\hat\nu_{i}|- \hat \omega_{i}}\]

where \(|\hat\nu_{i}|- \hat \omega_{i}\) is the prediction for the
conditional standard deviation
\(\sigma(x_i)=x_{i}' \gamma + \zeta_{g1} + \zeta_{g2}\)

The \(\tau_{th}\) quantile of the error \(\varepsilon\) can be estimated
as usual and the variance-covariance matrices obtained in the same way
as before (see Section~\ref{sec-se}) by using \(x_{i}^{rc}\) instead of
\(x_{i}\) when estimating the influence functions for all estimated
coefficients.

\hypertarget{simulation-evidence}{%
\section{Simulation Evidence}\label{simulation-evidence}}

To show the performance of the extended strategy, we implement a small
simulation study. We consider a simple model with a single explanatory
variable \(x\). In contrast with MSS, we consider a two-way fixed effect
structure that is more general than the panel structure. For this
exercise, we consider the following data-generating process:

\[y_i = \alpha_{1i} + \alpha_{2i} + x_i + (1+x_i + \alpha_{1i} + \alpha_{2i} ) \varepsilon_i
\]

where \(\alpha_{1i}\sim \chi^2(1)\), \(\alpha_{2i} \sim \chi^2(1)\), and
\(x_i = 0.5 * (\chi_i +0.5(\alpha_{1i}+\alpha_{2i}))\), with
\(\chi_i \sim \chi^2(1)\). We only consider the case when the error term
\(\varepsilon_i\) is assumed to be normally distributed with mean 0 and
standard deviation of 0.5. We assume that there are 50 mutually
exclusive fixed effect groups for each set of fixed effects.
Observations are assigned to each subgroup randomly using a uniform
distribution between 1 and 50. We consider sample sizes of 500, 1000,
2000, and 4000 observations, which implies an average of 10, 20, 40, and
80 observations per group. The model is estimated using the
location-scale model with heteroskedastic linear errors and we report
the coefficients for the 25th and 75th quantiles. We run this excercise
5000 times. Table~\ref{tbl-sim1} reports the bias, simulated standard
error, and mean squared error. We also report the results obtained using
an adaptation bias-corrected estimator based on the split-panel
jackknife estimator proposed by \citet{dhaene2015}.\footnote{For the
  implementation, we first estimate the model using the full sample,
  then randomly assign each observation into one of two groups, and
  finally reestimate the quantile coefficients for each group. The
  bias-corrected estimator is then obtained as
  \(\hat\beta(\tau)_{jkc}=2*\hat\beta(\tau)_{full}-0.5*\big(\hat\beta(\tau)_{s1}+\hat\beta(\tau)_{s1}\big)\).}
These results are labeled JKC.

Similar to the findings in MSS, we find that while the estimator
presents a bias when the sample is small (\(N=500\)), this bias shrinks
as the sample size increases. As MSS describes, the bias seems to be
proportional to the sample size, or more precisely to \(N/N_{g1}\) and
\(N/N_{g2}\), where \(N_{g1}\) and \(N_{g2}\) are the average number of
observations per group. Interestingly, the bias-corrected estimator
presents an almost 0 bias, even when the samples are small (about 10
observations per subgroup). Moreover, we find that despite the bias
reduction obtained using the JKC estimator, the standard errors and mean
squared error (MSE) are larger than without correction, which also
replicates the results of MSS,

\hypertarget{tbl-sim1}{}
\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\columnwidth - 14\tabcolsep) * \real{0.1250}}
  >{\centering\arraybackslash}p{(\columnwidth - 14\tabcolsep) * \real{0.1250}}
  >{\centering\arraybackslash}p{(\columnwidth - 14\tabcolsep) * \real{0.1250}}
  >{\centering\arraybackslash}p{(\columnwidth - 14\tabcolsep) * \real{0.1250}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 14\tabcolsep) * \real{0.1250}}
  >{\centering\arraybackslash}p{(\columnwidth - 14\tabcolsep) * \real{0.1250}}
  >{\centering\arraybackslash}p{(\columnwidth - 14\tabcolsep) * \real{0.1250}}
  >{\centering\arraybackslash}p{(\columnwidth - 14\tabcolsep) * \real{0.1250}}@{}}
\caption{\label{tbl-sim1}Bias, Simulated Standard error, and Mean
Squared Error}\tabularnewline
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\raggedright
N = 500
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
Bias
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
Sim SE
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
MSE
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
N = 1000
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
Bias
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
Sim SE
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
MSE
\end{minipage} \\
\midrule\noalign{}
\endfirsthead
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\raggedright
N = 500
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
Bias
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
Sim SE
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
MSE
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
N = 1000
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
Bias
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
Sim SE
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
MSE
\end{minipage} \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
q25 & & & & q25 & & & \\
~mmqreg & 0.104 & 0.247 & 0.072 & ~mmqreg & 0.050 & 0.173 & 0.032 \\
~jkc & 0.008 & 0.297 & 0.088 & ~jkc & -0.001 & 0.191 & 0.036 \\
q75 & & & & q75 & & & \\
~mmqreg & -0.091 & 0.251 & 0.071 & ~mmqreg & -0.044 & 0.176 & 0.033 \\
~jkc & 0.002 & 0.303 & 0.092 & ~jkc & 0.006 & 0.195 & 0.038 \\
\textbf{N = 2000} & \textbf{Bias} & \textbf{Sim SE} & \textbf{MSE} &
\textbf{N = 4000} & \textbf{Bias} & \textbf{Sim SE} & \textbf{MSE} \\
q25 & & & & q25 & & & \\
~mmqreg & 0.025 & 0.122 & 0.016 & ~mmqreg & 0.012 & 0.087 & 0.008 \\
~jkc & 0.003 & 0.129 & 0.017 & ~jkc & 0.000 & 0.089 & 0.008 \\
q75 & & & & q75 & & & \\
~mmqreg & -0.023 & 0.124 & 0.016 & ~mmqreg & -0.011 & 0.087 & 0.008 \\
~jkc & 0.001 & 0.131 & 0.017 & ~jkc & 0.000 & 0.089 & 0.008 \\
\end{longtable}

As described in section Section~\ref{sec-se}, GLS standard errors may
not be appropriate if heteroskedasticity in the model is misspecified.
This could happen if the heteroskedasticity is not linear in parameters
or if there are correlations within groups. To show the performance of
the GLS standard errors, we consider two additional simulations. To see
the impact of heteroskedasticity misspecification, we assume the DGP is
given by

\[y_i = \alpha_{1i} + \alpha_{2i} + x_i + (1+x_i + \sqrt{x_i} +\alpha_{1i} + \alpha_{2i} ) \varepsilon_i
\]

To show the impact of ignoring correlation within groups, we consider
the case where the error \(\varepsilon\) is correlated within groups
defined by the first fixed effect. Specifically:

\[\varepsilon_i =0.5*(\sqrt{0.5} z_i + \sqrt{0.5} w_g)\]

where \(z_i~N(0,1)\) and \(w_g~N(0,1)\), with \(g\) being the \(g_{th}\)
group defined by the first fixed effect.

For the simulation, we run the excercise 5000 times, using a sample of
size \(N=2000\). The results from both simulations are reported in the
Table~\ref{tbl-sim2}. We report the average coefficient,
simulation-based standard errors, and the standard errors using GLS,
robust and clustered estimations.

\hypertarget{tbl-sim2}{}
\begin{longtable}[]{@{}lcccc@{}}
\caption{\label{tbl-sim2}Heteroskedasticity misspecification and
clustered errors}\tabularnewline
\toprule\noalign{}
& Het:q25 & Het:q75 & Clust:q25 & Clust:q75 \\
\midrule\noalign{}
\endfirsthead
\toprule\noalign{}
& Het:q25 & Het:q75 & Clust:q25 & Clust:q75 \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
Coefficient & 0.550 & 1.442 & 0.775 & 1.229 \\
SIM\_SE & 0.154 & 0.155 & 0.109 & 0.107 \\
GLS\_SE & 0.476 & 0.451 & 0.288 & 0.221 \\
Robust & 0.140 & 0.139 & 0.087 & 0.088 \\
Cluster & & & 0.097 & 0.097 \\
\end{longtable}

One of the conclusions from the results in Table~\ref{tbl-sim2} is that
the results for the 25th and 75th quantiles are mostly identical,
because of the symmetric error we used in the simulation. On the one
hand, because the error structure is misspecified, we should expect to
see a large bias in the estimated coefficients, especially in cases of
misspecified heteroskedasticity. However, the purpose of this exercise
is not to demonstrate the unbiasedness of the coefficients, but rather
to assess the performance of the standard error estimation.

As expected, the GLS standard errors are biased and the bias is larger
in the case of misspecified heteroskedasticity. In fact, with
heteroskedasticity misspecification, the average GLS-SE (0.476) is
almost three times as large as the simulation-based standard error
(0.154). In this first case, robust standard errors perform better, but
with a small difference (0.14 to 0.154). In the case of clustered
correlation, GLS-SE also performs poorly. Robust standard errors perform
better, but seem to underestimate the magnitude of the true standard
errors. Finally, cluster standard errors perform the best, with a small
difference between the simulation-based standard errors (0.097 to
0.109).

\hypertarget{illustrative-application}{%
\section{Illustrative application}\label{illustrative-application}}

In this section we replicate one of the exercises from MSS, allowing for
time and individual fixed effects as well as for different standard
errors estimations. We use data from \citet{persson_economic_2005}, to
estimate the relationship between surplus of government as share of GDP,
and a measure of quality of democracy (POLITY); log of real income per
capita (LYP); trade volume as share of GDP (TRADE); share of population
between 15 and 65 years of age (P1564); the share of the population 65
years and older (P65); one-year lag of the dependent variable (LSP); oil
prices in US dollars, differntiating between importer and exporter
countries (OILIM and OILEX); and the output gap (YGAP). In addition to
country fixed effects (as illustrated in MSS), we show results allowing
for time fixed effects. Table~\ref{tbl-tb1} and Table~\ref{tbl-tb2}
provide the results for the model with and without time fixed effects,
respectively. The tables showcase the location and scale coefficients,
as well as the quantile coefficients for the 25th, 50th and 75th
quantiles. We also report GLS standard errors, robust standard errors
(brackets) and clustered standard errors at the country level.

\hypertarget{tbl-tb1}{}
\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\columnwidth - 18\tabcolsep) * \real{0.1235}}
  >{\centering\arraybackslash}p{(\columnwidth - 18\tabcolsep) * \real{0.1111}}
  >{\centering\arraybackslash}p{(\columnwidth - 18\tabcolsep) * \real{0.0988}}
  >{\centering\arraybackslash}p{(\columnwidth - 18\tabcolsep) * \real{0.0864}}
  >{\centering\arraybackslash}p{(\columnwidth - 18\tabcolsep) * \real{0.0864}}
  >{\centering\arraybackslash}p{(\columnwidth - 18\tabcolsep) * \real{0.0988}}
  >{\centering\arraybackslash}p{(\columnwidth - 18\tabcolsep) * \real{0.0988}}
  >{\centering\arraybackslash}p{(\columnwidth - 18\tabcolsep) * \real{0.0988}}
  >{\centering\arraybackslash}p{(\columnwidth - 18\tabcolsep) * \real{0.0988}}
  >{\centering\arraybackslash}p{(\columnwidth - 18\tabcolsep) * \real{0.0988}}@{}}
\caption{\label{tbl-tb1}Determinants of Government Surpluses: Individual
Fixed Effects}\tabularnewline
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
polityt
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
lyp
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
trade
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
p1564
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
p65
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
lspl
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
oil\_im
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
oil\_ex
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
ygap
\end{minipage} \\
\midrule\noalign{}
\endfirsthead
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
polityt
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
lyp
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
trade
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
p1564
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
p65
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
lspl
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
oil\_im
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
oil\_ex
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
ygap
\end{minipage} \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
Location & & & & & & & & & \\
~coeff & 0.116 & -0.715 & 0.030 & 0.121 & 0.028 & 0.691 & -0.047 &
-0.006 & 0.010 \\
~se\_gls & 0.046 & 0.540 & 0.008 & 0.033 & 0.070 & 0.035 & 0.008 & 0.022
& 0.028 \\
~se\_r & 0.047 & 0.597 & 0.008 & 0.031 & 0.070 & 0.037 & 0.007 & 0.017 &
0.021 \\
~se\_cl & 0.046 & 0.465 & 0.007 & 0.032 & 0.071 & 0.035 & 0.010 & 0.020
& 0.023 \\
Scale & & & & & & & & & \\
~coeff & -0.097 & -0.616 & 0.003 & 0.036 & 0.087 & -0.085 & 0.013 &
0.016 & -0.004 \\
~se\_gls & 0.032 & 0.371 & 0.005 & 0.023 & 0.048 & 0.024 & 0.006 & 0.015
& 0.019 \\
~se\_r & 0.031 & 0.398 & 0.005 & 0.020 & 0.049 & 0.025 & 0.005 & 0.010 &
0.015 \\
~se\_cl & 0.048 & 0.800 & 0.008 & 0.031 & 0.067 & 0.029 & 0.004 & 0.010
& 0.012 \\
Q25 & & & & & & & & & \\
~coeff & 0.191 & -0.239 & 0.028 & 0.093 & -0.039 & 0.756 & -0.057 &
-0.018 & 0.013 \\
~se\_gls & 0.059 & 0.684 & 0.010 & 0.042 & 0.088 & 0.045 & 0.010 & 0.027
& 0.035 \\
~se\_r & 0.056 & 0.656 & 0.008 & 0.036 & 0.086 & 0.040 & 0.010 & 0.020 &
0.025 \\
~se\_cl & 0.073 & 0.687 & 0.006 & 0.041 & 0.098 & 0.023 & 0.010 & 0.021
& 0.029 \\
Q50 & & & & & & & & & \\
~coeff & 0.108 & -0.765 & 0.030 & 0.124 & 0.035 & 0.684 & -0.046 &
-0.005 & 0.009 \\
~se\_gls & 0.046 & 0.535 & 0.007 & 0.033 & 0.069 & 0.035 & 0.008 & 0.022
& 0.027 \\
~se\_r & 0.046 & 0.593 & 0.008 & 0.031 & 0.069 & 0.036 & 0.007 & 0.017 &
0.021 \\
~se\_cl & 0.043 & 0.484 & 0.008 & 0.032 & 0.070 & 0.036 & 0.010 & 0.020
& 0.023 \\
Q75 & & & & & & & & & \\
~coeff & 0.031 & -1.258 & 0.033 & 0.153 & 0.104 & 0.616 & -0.036 & 0.008
& 0.006 \\
~se\_gls & 0.048 & 0.551 & 0.008 & 0.034 & 0.071 & 0.036 & 0.008 & 0.022
& 0.028 \\
~se\_r & 0.049 & 0.696 & 0.009 & 0.034 & 0.075 & 0.043 & 0.007 & 0.018 &
0.023 \\
~se\_cl & 0.039 & 0.919 & 0.012 & 0.041 & 0.079 & 0.055 & 0.010 & 0.022
& 0.020 \\
\end{longtable}

As expected, Table~\ref{tbl-tb1} shows that the point estimates are
identical to those reported in \citet{mss2019} (Table 6), including
analytical standard errors (GLS). With our estimator, however, we are
also able to produce both robust and clustered standard errors for
location and scale coefficients. Except for a few cases, the robust and
clustered standard errors are larger than the GLS standard errors, which
may be an indication of misspecification of the model. The GLS standard
errors we report differ from those in MSS, because they use panel
standard errors, which are equivalent to our clustered standard errors,
instead of the analytical standard errors we derive.

Considering the estimated effects across quantiles, we observe few
differences in the reported GLS standard errors compared to the
analytical standard errors reported in MSS. Our clustered standard
errors, however, are closer to the bootstrap-based standard errors the
authors report.\footnote{The differences in the GLS standard errors may
  be due that in our derivation the influence function of the
  standardized \(\tau_{th}\) quantile (see \#eq-infs) does not have the
  same leading term as the one reported in MSS (see Theorem 3, and the
  definition of \(W\)).}

In Table~\ref{tbl-tb2}, we report the results including both individual
and year fixed effects. Because oil prices only vary across years, the
variable is excluded from the model specification. Accounting for time
fixed effects does not change the general conclusions that can be drawn,
based on the results from Table~\ref{tbl-tb1}. The two largest
differences are that the log of income per capita has a positive effect
on government surpluses, but only for the 25th quantile, because at this
point the largest impact on the Scale component is felt. Similarly, we
observe that the income gap has an impact on government surplus that is
always negative and increasing across quantiles. In both instances, the
effects are not statistically significant.

\hypertarget{tbl-tb2}{}
\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\columnwidth - 14\tabcolsep) * \real{0.1493}}
  >{\centering\arraybackslash}p{(\columnwidth - 14\tabcolsep) * \real{0.1194}}
  >{\centering\arraybackslash}p{(\columnwidth - 14\tabcolsep) * \real{0.1194}}
  >{\centering\arraybackslash}p{(\columnwidth - 14\tabcolsep) * \real{0.1045}}
  >{\centering\arraybackslash}p{(\columnwidth - 14\tabcolsep) * \real{0.1493}}
  >{\centering\arraybackslash}p{(\columnwidth - 14\tabcolsep) * \real{0.1194}}
  >{\centering\arraybackslash}p{(\columnwidth - 14\tabcolsep) * \real{0.1194}}
  >{\centering\arraybackslash}p{(\columnwidth - 14\tabcolsep) * \real{0.1194}}@{}}
\caption{\label{tbl-tb2}The Determinants of Government Surpluses:
Individual and Time Fixed Effects}\tabularnewline
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
polity
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
lyp
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
trade
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
prop1564
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
prop65
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
lspl
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
ygap
\end{minipage} \\
\midrule\noalign{}
\endfirsthead
\toprule\noalign{}
\begin{minipage}[b]{\linewidth}\raggedright
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
polity
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
lyp
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
trade
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
prop1564
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
prop65
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
lspl
\end{minipage} & \begin{minipage}[b]{\linewidth}\centering
ygap
\end{minipage} \\
\midrule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
Location & & & & & & & \\
~coeff & 0.126 & -0.418 & 0.028 & 0.108 & 0.042 & 0.693 & -0.014 \\
~se\_gls & 0.087 & 1.157 & 0.015 & 0.072 & 0.136 & 0.066 & 0.053 \\
~se\_r & 0.047 & 0.703 & 0.008 & 0.038 & 0.068 & 0.038 & 0.022 \\
~se\_cl & 0.048 & 0.506 & 0.008 & 0.044 & 0.077 & 0.037 & 0.022 \\
Scale & & & & & & & \\
~coeff & -0.095 & -1.255 & 0.005 & 0.033 & 0.040 & -0.081 & 0.008 \\
~se\_gls & 0.081 & 1.073 & 0.014 & 0.067 & 0.126 & 0.061 & 0.049 \\
~se\_r & 0.031 & 0.452 & 0.005 & 0.025 & 0.045 & 0.025 & 0.017 \\
~se\_cl & 0.041 & 0.848 & 0.006 & 0.030 & 0.048 & 0.033 & 0.013 \\
Q25 & & & & & & & \\
~coeff & 0.201 & 0.576 & 0.025 & 0.082 & 0.010 & 0.757 & -0.020 \\
~se\_gls & 0.154 & 2.070 & 0.024 & 0.118 & 0.219 & 0.121 & 0.085 \\
~se\_r & 0.058 & 0.751 & 0.008 & 0.049 & 0.080 & 0.040 & 0.026 \\
~se\_cl & 0.073 & 0.761 & 0.006 & 0.052 & 0.087 & 0.023 & 0.027 \\
Q50 & & & & & & & \\
~coeff & 0.119 & -0.512 & 0.029 & 0.111 & 0.045 & 0.687 & -0.013 \\
~se\_gls & 0.091 & 1.230 & 0.014 & 0.070 & 0.130 & 0.072 & 0.051 \\
~se\_r & 0.046 & 0.695 & 0.008 & 0.037 & 0.068 & 0.038 & 0.022 \\
~se\_cl & 0.045 & 0.529 & 0.008 & 0.044 & 0.077 & 0.039 & 0.021 \\
Q75 & & & & & & & \\
~coeff & 0.041 & -1.555 & 0.033 & 0.138 & 0.078 & 0.619 & -0.007 \\
~se\_gls & 0.067 & 0.898 & 0.011 & 0.053 & 0.098 & 0.052 & 0.038 \\
~se\_r & 0.048 & 0.827 & 0.009 & 0.037 & 0.075 & 0.046 & 0.026 \\
~se\_cl & 0.038 & 0.980 & 0.012 & 0.050 & 0.086 & 0.063 & 0.020 \\
\end{longtable}

\hypertarget{conclusions}{%
\section{Conclusions}\label{conclusions}}

In this paper, we have extended the methodology proposed by
\citet{mss2019} in order to estimate quantile regression models with
multiple sets of fixed effects as well as with alternative standard
errrors. This methodology will allow researchers to implement
more-comprehensive analyses of data sets characterized by complex
hierarchies and unobserved heterogeneity. This extension is particularly
valuable in contexts where group specific effects vary across the
conditional distribution of the outcome of interest.

Using a small simulation study, we show that our extended approach is as
effective in identifying the parameters of interest as that of
\citet{mss2019}, even in contexts with two sets of fixed effects.
Notably, the bias-corrected estimator based on the split-panel jackknife
estimator exhibits promising results, mitigating biases when samples are
small but increasing standard errors.

Furthermore, we have assessed the impact of misspecified
heteroskedasticity and intracluster correlation on the performance of
standard error estimations. Our findings emphasize the importance of
using appropriate standard error estimators to ensure accurate
inference. In particular, we find that GLS standard errors are biased
when the model heteroskedasticity is misspecified and that robust and
clustered standard errors perform better in these cases. This may also
suggest a venue for future research, in which the validity of the
location-scale model assumption is tested.

\hypertarget{references}{%
\section*{References}\label{references}}
\addcontentsline{toc}{section}{References}

\renewcommand{\bibsection}{}
\bibliography{bibliography.bib}

\newpage{}

\appendix

\hypertarget{derivation-of-the-influence-functions}{%
\section{Derivation of the influence
functions}\label{derivation-of-the-influence-functions}}

\hypertarget{model-identification}{%
\subsection{Model Identification}\label{model-identification}}

The estimation of quantile regression via moments assumes that the DGP
is linear in parameters, with an heteroskedastic error term that is also
a linear function of parameters:

\[\begin{aligned}
y_i = x_i'\beta + \nu_i \\
\nu_i = \varepsilon_i \times x_i'\gamma
\end{aligned}
\]

where \(\varepsilon\) is an unobserved i.i.d. random variable that is
independent of \(x\) and such that \(x\gamma\) is larger than 0 for any
\(x\).

In this case, the \(\tau_{th}\) conditional quantile model can be
written as

\[q(y | \tau, x) = x'(\beta + q(\varepsilon | \tau) \times \gamma)\]

This model is identified under the following conditions:

\[\begin{aligned}
  E[(y_i-x_i'\beta)x_i ]  &= E[h_{1,i}]=0 \\
  E[ (|y_i-x_i'\beta|-x_i' \gamma) x_i ] &=E[h_{2,i}]=0 \\
  E\left[  \mathbb{1}\left( q(\tau) x_i'\gamma +x_i'\beta\geq  y_i  \right) - \tau \right] 
   &=E[h_{3,i}]=0 
\end{aligned}
\]

For simplicity, the rest of the appendix uses \(q^\varepsilon_\tau\) to
represent \(q(\varepsilon | \tau)\).

\hypertarget{estimation-of-the-variance-covariance-matrix}{%
\subsection{Estimation of the Variance-Covariance
Matrix}\label{estimation-of-the-variance-covariance-matrix}}

In this model, to estimate the variance-covariance matrix the set of
coefficients \(\theta'=[\beta' \ \gamma' \ q^\varepsilon_\tau]\), we
need to obtain the influence functions of all coefficients, which are
defined as

\[\lambda_i = \bar G(\theta)^{-1}
\begin{bmatrix}
h_{1,i} \\
h_{2,i} \\
h_{3,i}
\end{bmatrix}
\]

where the Jacobian matrix \(\bar G(\theta)\) is defined as

\[\bar G(\theta) = \begin{bmatrix}
\bar G_{11} & \bar G_{12} & \bar G_{13} \\
\bar G_{21} & \bar G_{22} & \bar G_{23} \\
\bar G_{31} & \bar G_{32} & \bar G_{13} \\
\end{bmatrix}
\]

with

\[\bar G_{j,k} = - \frac 1 N \sum_{i=1}^N \frac{\partial h_{j,i}}{\partial \theta_k'} \ \forall j,k \in 1,2,3
\]

First Moment Condition: Location Model

\[h_{1,i}=x_i(y_i-x_i'\beta)\]

\[\begin{aligned}
\bar G_{1,1} &=- \frac{1}{N} \sum_{i=1}^N \frac{\partial h_{1,i}}{\partial \beta'} \\
             &=- \frac{1}{N} \sum_{i=1}^N (-x_i x_i') \\
             &= N^{-1} X'X
\end{aligned}
\]

\[
\bar G_{1,2} = \bar G_{1,3} = 0
\]

Second Moment Condition: Scale model

\[h_{2,i}=x_i(|y_i-x_i'\beta|-x_i'\gamma)\]

\[\begin{aligned}
\bar G_{2,1} &= -\frac{1}{N} \sum \frac{\partial h_{2,i}}{\partial \beta'} \\
             &=  \frac{1}{N} \sum x_i x_i' \frac{y_i-x_i'\beta}{|y_i-x_i'\beta|} \\
\frac{y_i-x_i'\beta}{|y_i-x_i'\beta|} &= sign(y_i-x_i'\beta) \\            
\end{aligned}
\]

Under the assumption \(\varepsilon_i \times x\gamma\), or in this case
\(y_i-x_i'\beta\), is uncorrelated with \(x\), we can simplify the
expression as

\[\begin{aligned}
\bar G_{2,1} &= N^{-1} \left(N^{-1}\sum sign(y_i-x_i'\beta)\right) \sum x_i x_i' \\
&= N^{-1} E[sign(y_i-x_i'\beta)] X'X
\end{aligned}
\]

\[\begin{aligned}
\bar G_{2,2} &= -\frac{1}{N} \sum \frac{\partial h_{2,i}}{\partial \gamma'} \\
             &=  \frac{1}{N} \sum x_i x_i' \\
             &= N^{-1} X'X
\end{aligned}
\]

\[\bar G_{2,3}=0\]

Third Moment Condition: Quantile of Standardized Residual

\[\begin{aligned}
h_{3,i} &= \mathbb{1}\left(q(\tau) x_i'\gamma +x_i'\beta - y_i \geq 0 \right) - \tau \text{ or}\\
h_{3,i} &= \mathbb{1}\left( q(\tau)  \geq \frac{y_i- x_i'\beta}{x_i'\gamma}  \right) - \tau = \mathbb{1}\big( q(\tau)  \geq \varepsilon  \big) - \tau \\
\end{aligned}
\]

Because the indicator function \(\mathbb{1}()\) is not differentiable,
we borrow from the nonparametric literature and approximate the
indicator function with the integral of a kernel density function
\(I()\). This function, \(I()\), is a monotonic and symmetrical function
around 0, with a domain over the real numbers, and a range between 0 and
1.

With an arbirarily small bandwidth \(h\), this function will approximate
the indicator function:

\[
\lim_{h\rightarrow 0} I\left(\frac{z}{h}\right) \approx \mathbb{1}(z\geq 0)
\]

Thus the function \(h_{3,i}\) can be approximated as

\[h_{3,i} \approx I\left( q(\tau) x_i'\gamma +x_i'\beta - y_i  \right) - \tau\]

Now, we can obtain the Jacobian matrix \(\bar G_{3,1}\) as:

\[\begin{aligned}
\bar G_{3,1} &= -\frac{1}{N} \sum \frac{\partial h_{3,i}}{\partial \beta'} \\
             &= -N^{-1} \sum K_h(q(\tau) x_i'\gamma +x_i'\beta - y_i ) x_i'
\end{aligned}
\]

Under the assumption that we have enough observations within each
combination of \(x\) that are of multiplicative heteroskedasticity, we
have

\[\begin{aligned}
E(K_h(q(\tau) x_i'\gamma +x_i'\beta - y_i )|X) &= f_{y|X}(q(\tau) x_i'\gamma +x_i'\beta) \\
&=\frac{1}{x_i'\gamma} f_{\varepsilon}(q(\tau)) \\
\end{aligned}
\]

where \(f_{y|X}\) is the conditional probability density function of
\(y\) given \(X\) and \(f_\varepsilon\) is the unconditional
distribution of the standardized error. With this, we can rewrite the
Jacobian matrix as

\[\bar G_{3,1} = -N^{-1} f_{\varepsilon}\big(q(\tau)\big) \sum  \frac{x_i'}{x_i'\gamma} 
\]

Asymptotically, however, the expression \(\sum\frac{a_i}{b_i}\) can be
approximated using Taylor expansions by \(N\frac{\bar a}{\bar b}\).
\footnote{This approximation will be useful when we consider the
  estimation of the influence functions.} Thus, we can rewrite the last
term as

\[\bar G_{3,1} = - f_{\varepsilon}\big(q(\tau)) \frac{\bar x_i'}{\bar x_i'\gamma} 
\]

The Jacobian for the second matrix \(\bar G_{3,2}\) can be derived
similarly:

\[\begin{aligned}
\bar G_{3,2} &= -\frac{1}{N} \sum \frac{\partial h_{3,i}}{\partial \gamma'} \\
             &= -N^{-1} \sum K_h(q(\tau) x_i'\gamma +x_i'\beta - y_i ) q(\tau) x_i' \\
             &=-N^{-1} \sum f_{y|x}(q(\tau) x_i'\gamma +x_i'\beta ) q(\tau) x_i' \\
             &=-N^{-1} f_{\varepsilon}(q(\tau)) q(\tau) \sum \frac{x_i'}{x_i'\gamma} \\
             &=- f_{\varepsilon}(q(\tau)) q(\tau) \frac{\bar x_i'}{\bar x_i'\gamma}
\end{aligned}
\]

and the Jacobian for the third matrix \(\bar G_{3,3}\) is

\[\begin{aligned}
\bar G_{3,3} &= -\frac{1}{N} \sum \frac{\partial h_{3,i}}{\partial q(\tau)} \\
 &= -\frac{1}{N} \sum K_h(q(\tau) x_i'\gamma +x_i'\beta - y_i ) x_i'\gamma \\
 &= -\frac{1}{N} \sum f_{y|X}(q(\tau) x_i'\gamma +x_i'\beta) x_i'\gamma \\
 &= -\frac{1}{N} \sum f_{\varepsilon}(q(\tau)) \frac{x_i'\gamma}{ x_i'\gamma} \\
 &= - f_{\varepsilon}(q(\tau)) 
\end{aligned}
\]

\hypertarget{influence-functions}{%
\subsection{Influence Functions}\label{influence-functions}}

Location Coefficients

\[\lambda_i(\beta) = \bar G^{-1}_{1,1} h_{1,i} = N (X'X)^{-1}(x_i (y_i-x_i'\beta)) = N (X'X)^{-1}(x_i \nu_i)
\]

which can also be written as a function of the standardized residuals:

\[\lambda_i(\beta) = \bar G^{-1}_{1,1} h_{1,i} = N (X'X)^{-1}(x_i (y_i-x_i'\beta)) = N (X'X)^{-1}(x_i ( x_i'\gamma \times \varepsilon))
\]

Scale Coefficients

\[\begin{aligned}
\lambda_i(\gamma)&=\bar G_{2,2}^{-1}\Big(h_{2,i}-\bar G_{2,1} \lambda_i(\beta)\Big) \\
&=N (X'X)^{-1} \Big(x_i(|\nu_i|-x_i' \gamma) 
- N^{-1} E[sign(\nu_i)] X'X \big[ N (X'X)^{-1}(x_i \nu_i) \big] \Big) \\
&=N (X'X)^{-1} \Big(x_i(|\nu_i|-x_i' \gamma) - E[sign(\nu_i)] (x_i \nu_i)  \Big) \\
&=N (X'X)^{-1} x_i \Big( |\nu_i| - E[sign(\nu_i)] \nu_i -x_i' \gamma \Big) 
\end{aligned}
\]

However,

\[\begin{aligned}
|\nu_i| &= \nu_i \times \mathbb{1}(\nu_i \geq 0) - \nu_i \times \mathbb{1}(\nu_i < 0) \\
|\nu_i| &= \nu_i \times \mathbb{1}(\nu_i \geq 0) - \nu_i \times [1-\mathbb{1}(\nu_i \geq 0)] \\
|\nu_i| &= 2 \nu_i \times \mathbb{1}(\nu_i \geq 0) - \nu_i  
\end{aligned}
\]

and

\[\begin{aligned}
E[sign(\nu_i)] &= E[\mathbb{1}(\nu_i \geq 0)] - E[\mathbb{1}(\nu_i < 0)] \\
E[sign(\nu_i)] &= E[\mathbb{1}(\nu_i \geq 0)] - E[(1-\mathbb{1}(\nu_i \geq 0))] \\
E[sign(\nu_i)] &= 2 E[\mathbb{1}(\nu_i \geq 0)] - 1
\end{aligned}
\]

Thus,

\[\begin{aligned}
\lambda_i(\gamma) &= N(X'X)^{-1} x_i   \Big( 2 \nu_i \times \mathbb{1}(\nu_i \geq 0) - \nu_i  - ( 2 E[\mathbb{1}(\nu_i \geq 0)] - 1) \nu_i -x_i' \gamma \Big) \\
  &= N(X'X)^{-1} x_i   \Big( 2 \nu_i \times \mathbb{1}(\nu_i \geq 0) -  2 E[\mathbb{1}(\nu_i \geq 0)] \nu_i -x_i' \gamma \Big) \\
  &= N(X'X)^{-1} x_i   \Big( 2 \nu_i \times \big[ \mathbb{1}(\nu_i \geq 0) -  E[\mathbb{1}(\nu_i \geq 0)] \big] -x_i'\gamma \Big) \\
  &= N(X'X)^{-1} x_i   \Big( \tilde \nu_i -x_i' 
  \gamma \Big)
\end{aligned}
\]

This last expression is the equivalent simplification used in
\citet{mss2019} and \citet{im2000}. If the scale function is strictly
positive, it also follows that
\(\mathbb{1}(\nu_i \geq 0)= \mathbb{1}(\varepsilon_i \geq 0)\). Thus, it
can be simplified as

\[\begin{aligned}
\lambda_i(\gamma) &= N(X'X)^{-1} x_i ( x_i' \gamma ) \times (\tilde \varepsilon_i -1\big)
\end{aligned}
\]

Quantile of Standardized Residual

\[\begin{aligned}
\lambda_i(q(\tau))&=\bar G_{3,3}^{-1}
\Big(
 h_{3,i}-\bar G_{3,1} \lambda_i(\beta)-\bar G_{3,2} \lambda_i(\gamma)
\Big) \\
&=-\frac{1}{f_{\varepsilon}(q(\tau))} \times \Bigg( \Big(\mathbb{1} ( q(\tau)  \geq \varepsilon  ) - \tau \Big)  \\
&+ f_{\varepsilon} (q(\tau)) \frac{\bar x_i'}{\bar x_i'\gamma} 
N (X'X)^{-1} x_i  ( x_i'\gamma \times \varepsilon) \\
&+ f_{\varepsilon}(q(\tau)) q(\tau) \frac{\bar x_i'}{\bar x_i'\gamma} N(X'X)^{-1} x_i   \big( \tilde \nu_i -x_i' 
  \gamma \big) 
\Bigg) \\
&=\frac{\tau-\mathbb{1}\big( q(\tau)  \geq \varepsilon  \big) }{f_{\varepsilon}(q(\tau))}
- \frac{ x_i'\gamma \times \varepsilon_i }{\bar x_i'\gamma} 
-  q(\tau) \frac{ \tilde \nu_i -x_i' 
  \gamma }{\bar x_i'\gamma} 
\end{aligned}
\]

\hypertarget{implementation}{%
\section{Implementation}\label{implementation}}

The method described here can be implemented using any of the following
packages:

\begin{itemize}
\tightlist
\item
  \texttt{mmqreg} in Stata: `net install mmqreg,
  from(https://friosavila.github.io/stpackages)
\item
  \texttt{mmqreg} in R: Program available in mmqreg repository on GitHub
  (https://github.com/friosavila/mmqreg)
\item
  \texttt{mmqreg} in Python:
\item
  \texttt{mmqreg} in Julia:
\end{itemize}

\end{document}