### 2024-09-10 article/_extensions/sim-art2/_authors.tex
Update Paper
WIth Comments from GJ

$-- You can use as many custom partials as you need. Convention is to prefix name with '_'
$-- It can be useful to use such template to split some template parts in smaller pieces, which is easier to reuse. 
$-- This '_custom.tex' is used on 'title.tex' as example.
$-- See other existing format in quarto-journals/ organisation.
$-- %%%% TODO %%%%%
$-- Use it if you need to insert content at this specific place of the main Pandoc's template. Otherwise, remove it.
$-- Here we are using it to format the authors part of the template.
$-- %%%%%%%%%%%%%%%

$it.name.literal$
\thanks{
  $for(it.affiliations)$
    $it.name$$sep$, $it.city$
  $endfor$
  $if(it.email)$
    \href{mailto:$it.email$}{Email: $it.email$}
  $endif$
}