### 2024-09-10 article/alt_article.tex

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
\PassOptionsToPackage{dvipsnames,svgnames,x11names}{xcolor}
%
\documentclass[
  12pt,
  oneside]{article}

\usepackage{amsmath,amssymb}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
\usepackage{lmodern}
\ifPDFTeX\else  
    % xetex/luatex font selection
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\usepackage[lmargin=1.1in,rmargin=1.1in,tmargin=1in,bmargin=1.2in]{geometry}
\setlength{\emergencystretch}{3em} % prevent overfull lines
\setcounter{secnumdepth}{5}
% Make \paragraph and \subparagraph free-standing
\makeatletter
\ifx\paragraph\undefined\else
  \let\oldparagraph\paragraph
  \renewcommand{\paragraph}{
    \@ifstar
      \xxxParagraphStar
      \xxxParagraphNoStar
  }
  \newcommand{\xxxParagraphStar}[1]{\oldparagraph*{#1}\mbox{}}
  \newcommand{\xxxParagraphNoStar}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
  \let\oldsubparagraph\subparagraph
  \renewcommand{\subparagraph}{
    \@ifstar
      \xxxSubParagraphStar
      \xxxSubParagraphNoStar
  }
  \newcommand{\xxxSubParagraphStar}[1]{\oldsubparagraph*{#1}\mbox{}}
  \newcommand{\xxxSubParagraphNoStar}[1]{\oldsubparagraph{#1}\mbox{}}
\fi
\makeatother

\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother

% TODO: Add custom LaTeX header directives here
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{Figure}
\else
  \newcommand\figurename{Figure}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother

\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\usepackage[]{natbib}
\bibliographystyle{chicago}
\usepackage{bookmark}

\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={Estimating Quantile Regressions with Multiple Fixed Effects through Method of Moments},
  pdfauthor={Fernando Rios-Avila; Leonardo Siles; Gustavo Canavire-Bacarreza},
  pdfkeywords={Fixed effects, Linear heteroskedasticity, Location-scale
model, Quantile regression},
  colorlinks=true,
  linkcolor={blue},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue},
  pdfcreator={LaTeX via pandoc}}

\definecolor{ahjcolor}{rgb}{0.0, 0.13, 0.82}
\PassOptionsToPackage{colorlinks=true, linkcolor=ahjcolor, urlcolor=ahjcolor, citecolor=ahjcolor}{hyperref}

% --- Packages ---
% Layout and spacing
\usepackage{setspace}
\usepackage{enumitem}

% Math and symbols
\usepackage{amsmath, amssymb, amsthm, amsbsy, amsfonts, mathrsfs, mathtools, bm, dsfont}

% Graphics and tables
\usepackage{graphicx, float, caption, subcaption}
\usepackage{longtable, array, multirow, booktabs, threeparttable}
\usepackage{lscape}

% Text formatting
\usepackage{color, xcolor, soul}
\usepackage{titlesec}

% Additional features
\usepackage{authblk} % Author and affiliation
\usepackage{appendix} % Appendix support

% Hyperlinks
\usepackage{hyperref}

% Math Operators
\DeclareMathOperator*{\argmin}{arg\,min}
\DeclareMathOperator*{\argmax}{arg\,max}
\DeclareMathOperator*{\sign}{sign}
\DeclareMathOperator*{\tr}{Tr}
\DeclareMathOperator*{\p}{Pr}
\DeclareMathOperator*{\Var}{Var}
\DeclareMathOperator*{\rank}{rank}
\DeclareMathOperator*{\diag}{diag}

% Theorem Environments
\newtheorem{thm}{Theorem}
\newtheorem{cor}{Corollary}
\newtheorem{lemma}{Lemma}
\newtheorem{prop}{Proposition}
\newtheorem{ass}{Assumption}
\newtheorem{exm}{Example}

% Affiliation formatting
\renewcommand\Affilfont{\itshape\small}

% LaTeX tweaks
\setcounter{MaxMatrixCols}{10}
\makeatletter
\makeatother

% Page settings
\onehalfspacing
\setlength{\parskip}{0.5em}
\setlist{nolistsep}
% title

\title{\textbf{\Large Estimating Quantile Regressions with Multiple
Fixed Effects through Method of Moments}\thanks{The opinions expressed
in this paper are those of the authors and do not necessarily reflect
the views of the World Bank, its Board of Directors, or the countries it
represents.

The authors would like to thank Enrique Pinzon an Joao Santos-Silva for
their helpful comments and suggestions.}}

\author{
Fernando Rios-Avila
\thanks{
      Levy Economics Institute of Bard
College      \href{mailto:friosavi@levy.org}{Email: friosavi@levy.org}
  } \hspace{0.25 cm} 
Leonardo Siles
\thanks{
      Universidad Católica
Boliviana      \href{mailto:leonardo.siles@ucb.edu.bo}{Email: leonardo.siles@ucb.edu.bo}
  } \hspace{0.25 cm} 
Gustavo Canavire-Bacarreza
\thanks{
      The World
Bank      \href{mailto:gcanavire@worldbank.org}{Email: gcanavire@worldbank.org}
  }}

\begin{document}
\maketitle
\begin{abstract}

    This paper proposes a new method to estimate quantile regressions
with multiple fixed effects. The method, which expands on the strategy
proposed by \citet{mss2019}, allows for the inclusion of multiple fixed
effects and provides various alternatives for the estimation of standard
errors. We provide Monte Carlo simulations to show the finite sample
properties of the proposed method in the presence of two sets of fixed
effects. Finally, we apply the proposed method to estimate the
determinants of the surplus of government as a share of GDP, allowing
for both time and country fixed effects. \\ \\

    \noindent \textbf{Keywords:} Fixed effectsLinear
heteroskedasticityLocation-scale modelQuantile regression \\
    \noindent \textbf{JEL Codes:} C21C22C23

    \end{abstract}

    \newpage

\newpage{}

\section{Introduction}\label{introduction}

Quantile regression (QR), introduced by \citet{koenkerbasset1978}, is an
estimation strategy used for modeling the relationships between
explanatory variables X and the conditional quantiles of the dependent
variable \(Q_y (\tau|x)\). Using QR one can obtain richer
characterizations of the relationships between dependent and independent
variables, by exploring how the variables relate along the entire
conditional distribution.

A relatively recent development in the literature has focused on
extending quantile regressions analysis to include individual fixed
effects in the framework of panel data. However, as described in
\citet{neymanscott1948}, and \citet{lancaster2000}, when individual
fixed effects are included in quantile regression analysis an incidental
parameter problem is generated. While many strategies have been proposed
for estimating this type of model (see \citet{galvao2017quantile} for a
brief review), neither has become standard because of their restrictive
assumptions in regard to the inclusion of individual and multiple fixed
effects, the computational complexity, and implementation.

More recently, \citet{mss2019} (MSS hereafter) proposed a methodology
based on a conditional location-scale model, similar to the one
described in \citet{he1997} and \citet{zhao2000}, for the estimation of
quantile regressions models for panel data via a method of moments. This
method allows individual fixed effects to have heterogeneous effects on
the entire conditional distribution of the outcome, rather than
constraining their effect to be a location shift only, as in
\citet{canay2011}, \citet{koenker2004}, and \citet{lamarche2010}.

In principle, under the assumption that data-generating process behind
the data is based on a multiplicative heteroskedastic process that is
linear in parameters \citep[@he1997, @zhao2000, @cameron2005]{mss2019},
the effect of a variable \(X\) on the \(q_{th}\) quantile can be derived
as the combination of a location effect, and scale effect moderated by
the quantile of an underlying i.i.d. error. For statistical inference,
MSS derives the asymptotic distribution of the estimator, suggesting the
use of bootstrap standard errors, as well.

This methodology is not meant to substitute for the use of standard
quantile regression analysis. That said, given the assumptions required
for the identification of the model, it provides a simple and fast
alternative for the estimation of quantile regression models with
individual fixed effects.

In this framework, our paper expands on \citet{mss2019} in two ways.
First, making use of the properties of generalized method of moments
(GMM) estimators, we derive various alternatives for the estimation of
standard errors based on the empirical influence functions of the
estimators. Even if the model is correctly specified, robust standard
errors perform better than GLS standard errors due small violations of
the model assumptions due to sampling variability. Furthermore,
clustered standard errors may help to further account for typically
unobserved correlations across observations. Second, we reconsider the
application of Frisch-Waugh-Lovell (FWL) theorem \citep[ and
\citet{lovell1963}]{frishwaugh1933} to extend the MSS estimator and
allow for the inclusion of multiple fixed effects. This extension may be
useful for empirical analysis, as it is common to control for multiple
fixed effects such as individual and time fixed effects.

The rest of the paper is structured as follows: section 2 presents the
basic setup of the location-scale model described in \citet{he1997} and
\citet{zhao2000}, tying the relationship between the standard quantile
regression model and the location-scale model. It also revisits the
methodology of MSS, proposing alternative estimators for the standard
errors based on the properties of GMM estimators and the empirical
influence functions. It also shows that the FWL theorem can be used to
control for multiple fixed effects. Section 3 presents the results of a
small simulation study and section 4 illustrates the application of the
proposed methods with one empirical example. Section 5 concludes.

\section{Methodology}\label{sec-method}

\subsection{Quantile Regression: Location-Scale model}\label{sec-betas}

Quantile regressions are used to identify relationships between the
explanatory variables \(X\) and the conditional quantiles of the
dependent variable \(Q_y(\tau|X)\). This relationship is commonly
assumed to follow a linear functional form:

\begin{equation}\phantomsection\label{eq-eq1}{Q_y(\tau|X) =X\beta(\tau)
}\end{equation}

This allows for a linear effect of \(X\) on \(Y\), but that could vary
across values of \(\tau\).

An alternative formulation of quantile regressions is the location-scale
model. This approach assumes that the conditional quantile of \(Y\)
given \(X\) and \(\tau\) can be expressed as a combination of two
models: the location model, which describes the central tendency of the
conditional distribution; and the scale model, which describes
deviations from the central tendency:

\begin{equation}\phantomsection\label{eq-eq2}{Q_y(\tau|X) =X\beta+X\gamma(\tau)
}\end{equation}

Here, the location parameters \(\beta\) are typically identified using a
linear regression model (as in \citet{mss2019}) or a median regression
(as in \citet{he1997} and \citet{zhao2000}) and the scale parameters
\(\gamma(\tau)\) can be estimated using standard approaches.

Both the standard quantile regression (Equation~\ref{eq-eq1}) and the
location-scale specification (Equation~\ref{eq-eq2}) can be estimated as
the solution to a weighted minimization problem:

\begin{equation}\phantomsection\label{eq-eq3}{\hat{\beta}(\tau) = \underset{\beta}{\operatorname{argmin}}
\left( \sum_{i\in y_i\geq x_i'\beta} \tau (y_i - x_i'\beta) - \sum_{i\in y_i<x_i'\beta} (1-\tau)(y_i - x_i'\beta) \right)
}\end{equation}

One characteristic of this estimator is that the \(\beta(\tau)\)
coefficients are identified locally and thus the estimated quantile
coefficients will exhibit considerable variation when analyzed across
\(\tau\). It is also implicit that if one requires an analysis of the
entire distribution, it would be necessary to estimate the model for
each quantile.\footnote{There are other estimators that provide smoother
  estimates for the quantile regression coefficients using a kernel
  local weighted approach \citep{kaplan2017}, as well as identifying the
  full set of quantile coefficients while simultaneously assuming some
  parametric functional forms \citep{frumentobotai2016}.}

One insightful extension to the location-scale parameterizations
suggested by \citet{he1997}, \citet{zhao2000}, \citet{cameron2005}, and
\citet{mss2019} is to assume that the data-generating process (DGP) can
be written as a linear model with a multiplicative heteroskedastic
process that is linear in parameters.\footnote{\citet{mss2019} also
  discuss a model where heteroskedasticity can be an arbitrary nonlinear
  function \(\sigma(x_i'\gamma)\), but develop the estimator for the
  linear case, i.e., when \(\sigma()\) is the identity function.}

\begin{equation}\phantomsection\label{eq-eq4}{\begin{aligned}
y_i &=x_i'\beta+\nu_i \\
\nu_i &=\varepsilon_i \times x_i'\gamma 
\end{aligned}
}\end{equation}

Under the assumption that \(\varepsilon\) is an i.i.d. unobserved random
variable that is independent of \(X\), the conditional quantile of \(Y\)
given \(X\) and \(\tau\) can be written as

\begin{equation}\phantomsection\label{eq-eq5}{Q_y(\tau|X) =X\beta+Q_\varepsilon(\tau) \times X\gamma 
}\end{equation}

In this setup, the traditional quantile coefficients are identified as
the location model coefficients plus the scale model coefficients
moderated by the \(\tau_{th}\) unconditional quantile of the
standardized error \(\varepsilon\). For simplicity we will use
\(q_\tau\) to denote \(Q_\varepsilon(\tau)\) in the rest of the paper.

\begin{equation}\phantomsection\label{eq-eq6}{\beta(\tau) = \beta + q_\tau \times \gamma 
}\end{equation}

While this specification imposes a strong assumption on the DGP, it has
two advantages over the standard quantile regression model. First,
because the location-scale model can be identified globally, with only a
single parameter (\(q_\tau\)) requiring local estimation, this
estimation approach will be more efficient than the standard quantile
regression model \citep{zhao2000}. Second, under the assumption that
\(X\gamma\) is strictly positive, the model will produce quantile
coefficients that do not cross \citep{he1997}.

Following MSS, the quantile regression model defined by
Equation~\ref{eq-eq5} can be estimated using a generalized method of
moments approach. And while it is possible to identify all coefficients
(\(\beta,\gamma, q_\tau\)) simultaneously, we describe and use the
implementation approach advocated by MSS, which identifies each set of
coefficients separately.

First, the location model can be estimated using a standard linear
regression model, where the dependent variable is the outcome \(Y\) and
the independent variables are the explanatory variables \(X\) (including
a constant) with an error \(u\), which is by definition heteroskedastic.
In this case, the location-model coefficients are identified under the
following condition:

\begin{equation}\phantomsection\label{eq-eq7}{\begin{aligned}
      y_i &=x_i'\beta+\nu_i \\
      E\big[ x_i \nu_i\big] &=0
      \end{aligned}
}\end{equation}

Second, after the location model is estimated, the scale coefficients
can be identified by modeling heteroskedasticity as a linear function of
characteristics \(X\). For this we use the absolute value of the errors
from the location model \(u\) as dependent variable, which allows us to
estimate the conditional standard deviation (rather than conditional
variance) of the errors. In this case, the coefficients are identified
under the following condition:

\begin{equation}\phantomsection\label{eq-eq8}{\begin{aligned}
  |\nu_i| &=x_i'\gamma+\omega_i \\
   E\big[ x_i (|\nu_i| -x_i'\gamma ) \big] &=0
  \end{aligned}
}\end{equation}

It should be noticed that the estimation of the standard errors (next
section) requires that the Scale component prediction \(x_i'\gamma\) is
strictly positive, because it represents the conditional standard
deviation of the error \(\nu_i\). Because this component is identified
using a linear model, some values for \(x_i'\gamma\) may be negative,
which will affect the estimation of the standard errors, as shown in the
simulation study.

Third, given the location and scale coefficients, the \(\tau_{th}\)
quantile of the error \(\varepsilon\) can be estimated using the
following condition:

\begin{equation}\phantomsection\label{eq-eq9}{\begin{aligned}
  E\left[  \mathbb{1}\left(x_i' (\beta +   \gamma q_\tau) \geq y_i \right) - \tau \right] &=0  \\
  E\left[  \mathbb{1}\left(   q_\tau \geq \frac{y_i-x_i'\beta}{x_i'\gamma} \right) - \tau \right] &=0  \\
  \end{aligned}
}\end{equation}

where one identifies the quantile of the error \(\varepsilon\) using
standardized errors \(\frac{y_i-x_i'\beta}{x_i'\gamma}\) or by finding
the values that identify the overall quantile coefficients
\(\beta(\tau)=\beta + \gamma q_\tau\). Afterwards, the conditional
quantile coefficients are simply defined as the combination of the
location and scale coefficients.

\subsection{Standard Errors: GLS, Robust, Clustered}\label{sec-se}

As discussed in the previous section, the estimation of quantile
regression coefficients using the location-scale model with
heteroskedastic linear errors can be estimated using a the following set
of moments, which fits within the GMM framework:

\begin{equation}\phantomsection\label{eq-eq10}{\begin{aligned}
  E[x_i \nu_i  ] &= E[h_{1,i}]=0 \\
  E[x_i  (|\nu_i|-x_i \gamma) ] &=E[h_{2,i}]=0 \\
  E\left[  \mathbb{1}\left(   q_\tau\geq \frac{y_i-x_i'\beta}{x_i'\gamma} \right) - \tau \right]  &=E[h_{3,i}]=0 
  \end{aligned}
}\end{equation}

Under the conditions described in \citet{newey_chapter_1994} (see
section 7), \citet{cameron2005} (see chapter 6.3.9), or as shown in
\citet{mss2019}, the location, scale, and residual quantile coefficients
are asymptotically normal.\footnote{\citet{zhao2000} also shows that the
  quantile coefficients for the location-scale model follow a normal
  distribution, but uses the assumption that the location model is
  derived using a least absolute deviation approach (median regression).}

Call \(\theta=[ \beta' \ \ \gamma' \ \ q_\tau ]'\) the set of
coefficients that are identified by the moment conditions in
Equation~\ref{eq-eq10}, a just identified model, and the function
\(h_i\) is a vector function that stacks all the moments described in
Equation~\ref{eq-eq10} at the individual level. Then \(\hat\theta\)
follows a normal distribution with mean \(\theta\) and
variance-covariance matrix \(V(\theta)\) that is estimated as

\[
\hat{V}(\hat\theta)=\frac{1}{N} 
\bar G(\hat\theta)^{-1} 
\left( \frac{1}{N} \sum_{i=1}^N h_i h_i'  \Big|_{\theta=\hat\theta} \right) 
\bar G(\hat\theta)^{-1} 
\]

which is equivalent to the Eicker-White heteroskedasticity-consistent
estimator for least-squares estimators.

Here, the inner product is the moment covariance matrix and
\(\bar{G}(\theta)\) is the Jacobian matrix of the moment equations
evaluated at \(\hat\theta\).

\[\bar{G}(\theta) =-\frac{1}{N} \sum_{i=1} \frac{\partial h_i}{\partial \theta'} \Big|_{\theta=\hat\theta}\]

In this framework, the quantile regression coefficients, a combination
of the location-scale-quantile estimates, will follow a normal
distribution with mean \(\beta(\tau) = \beta+\gamma q_\tau\) and
variance-covariance matrix equal to

\[\hat{V}(\beta(\tau)) = \Xi \hat{V}(\hat\theta) \Xi'
\]

where \(\Xi\) is a \(k \times (2k+1)\) matrix defined as

\begin{equation}\phantomsection\label{eq-eqqtile}{\Xi = [ I(k), \hat q_\tau \times I(k), \hat \gamma ]
}\end{equation}

with \(I(k)\) being an identity matrix of dimension \(k\) (number of
explanatory variables in \(X\) including the constant).

While it is possible to estimate the variance-covariance matrix using
simultaneous model estimation for a just identified model, it is more
efficient to estimate each set of coefficients separately. Afterward,
the variance-covariance matrix can be estimated using the empirical
influence functions of the estimators (see \citet{jann_2020} for an
overview of the application and \citet{hampel2005} for an in-depth
review).

Specifically, given an arbitrary vector of empirical influence functions
\(\lambda_i(\theta)\), the variance-covariance matrix can be estimated
as

\begin{equation}\phantomsection\label{eq-eqvcv}{\hat{V}(\theta) = \frac{1}{N^2} \sum_{i=1}^N \lambda_i(\theta) \lambda_i(\theta)'
}\end{equation}

where the influence functions are defined as:

\[\lambda_i(\theta) = \bar G(\theta)^{-1} h_i(\theta)\]

For the specific case of quantile regressions via moments, the influence
functions for the location, scale, and quantile coefficients
are\footnote{The derivation of the influence functions can be found in
  the appendix.}

\begin{equation}\phantomsection\label{eq-ifs}{\begin{aligned}
\lambda_i(\theta)&=
  \begin{bmatrix}
  \lambda_{i}(\beta) \\
  \lambda_{i}(\gamma) \\
  \lambda_{i}(q_\tau) 
  \end{bmatrix} \\
\lambda_{i}(\beta)&=N (X'X)^{-1}  x_i ( x_i'\gamma) \times \varepsilon_i \\
\lambda_{i}(\gamma)&= N(X'X)^{-1} x_i ( x_i' \gamma ) \times (\tilde \varepsilon_i -1\big)\\
\lambda_{i}(q_\tau)&=\frac{\tau-\mathbb{1}\big( q_\tau  \geq \varepsilon_i  \big) }{f_{\varepsilon}(q_\tau)}
- \frac{ x_i'\gamma \times \varepsilon_i }{\bar x_i'\gamma} 
-  q_\tau \frac{  x_i' \gamma  \times (\tilde \varepsilon_i -1\big)}{\bar x_i'\gamma}
\end{aligned}
}\end{equation}

The different types of standard errors estimation thus depend on the
assumptions imposed for the estimation of \(V(\theta)\).

\subsubsection{Robust Standard Errors}\label{robust-standard-errors}

The first and most natural standard error estimator is given by equation
Equation~\ref{eq-eqvcv}. This is equivalent to the Eicker-White
heteroskedasticity-consistent estimator for least-squares estimators.
Considering the location-scale model, the variance-covariance matrix for
the quantile coefficients can be estimated as

\[\hat{V}_{robust}
  \begin{pmatrix}
  \hat\beta \\
  \hat\gamma \\
  \hat q_\tau
  \end{pmatrix}
 = \frac{1}{N^2} 
 \begin{pmatrix}
 \sum \lambda_i(\beta) \lambda_i(\beta)' & \sum \lambda_i(\beta) \lambda_i(\gamma)' & \sum \lambda_i(\beta) \lambda_i(q_\tau)' \\
 \sum \lambda_i(\gamma) \lambda_i(\beta)' & \sum \lambda_i(\gamma) \lambda_i(\gamma)' & \sum \lambda_i(\gamma) \lambda_i(q_\tau)' \\
 \sum \lambda_i(q_\tau) \lambda_i(\beta)' & \sum \lambda_i(q_\tau)\lambda_i(\gamma)' & \sum \lambda_i(q_\tau) \lambda_i(q_\tau)' 
  \end{pmatrix}
\]

\subsubsection{Clustered Standard
Errors}\label{clustered-standard-errors}

Because one of the typical applications of quantile regressions is the
analysis of panel data, allowing for clustered standard errors at the
individual level is important. If the unobserved error \(\varepsilon\)
is correlated within clusters, generalized least squares (GLS) standard
errors could be severely biased. The standard recommendation has been to
report block-bootstrap standard errors, clustered at the individual
level.

Because we have access to the influence functions, it is straightforward
to estimate one-way clustered standard errors.

Let \(N_G\) be the total number of clusters \(g\), where
\(g=1\dots N_G\). The clustered variance covariance matrix is given
by\footnote{It should be noted that one could just as well apply the
  insights of \citet{cameron_robust_2011}, allowing for multiway
  clustering.}

\[\hat{V}_{clustered}
  \begin{pmatrix}
  \hat\beta \\
  \hat\gamma \\
  \hat q_\tau
  \end{pmatrix}
 = \frac{1}{N^2} 
 \begin{pmatrix}
 \sum_{g=1}^{N_G} S\lambda_i(\theta) S\lambda_i(\theta)'
  \end{pmatrix}
\]

where \(S\lambda_i(\theta)\) is the sum of the influence functions over
all observations within a given cluster \(g\).

\[S\lambda_i(\theta) = \sum_{i\in g} \lambda_i(\theta)
\]

\subsubsection{GLS Standard Errors}\label{sec-gls}

The standard errors proposed by MSS can be understood as an application
of GLS, which will be valid as long as the model for heteroskedasticity
is correctly specified.\footnote{As discussed in most econometric
  textbooks, for example \citet{cameron2005}, one approach to correct
  for heteroskedasticity, when the heteroskedasticity functional form is
  known or can be estimated, is to use weighted least squares. While
  feasible this approach would defeat the purpose of identifying
  quantile effects exploiting the heteroskedasticity of the model.} To
estimate the GLS standard errors, we make use of the following property:

Consider the influence functions and robust variance-covariance matrix
for the location coefficients:

\[\begin{aligned}
\hat V(\hat \beta) &= \frac{1}{N} \sum_i^N \lambda_{i}(\beta) \lambda_{i}(\beta)' \\
&= \frac{1}{N}  (X'X)^{-1}  \sum_i^N x_i x_i' ( x_i'\gamma \times \varepsilon_i)^2 (X'X)^{-1} \\
\end{aligned}
\]

Under the assumption that the model for heteroskedasticity is correctly
specified, we can apply the law of iterated expectations and rewrite the
variance-covariance matrix as

\[\begin{aligned}
\hat V(\hat \beta) &= \frac{1}{N} \sum_i^N \lambda_{i}(\beta) \lambda_{i}(\beta)' \\
&= E(\varepsilon_i^2) \frac{1}{N}  (X'X)^{-1}  \sum_i^N x_i x_i' ( x_i'\gamma)^2 (X'X)^{-1} \\
&= \sigma^2_{\varepsilon} \frac{1}{N}  (X'X)^{-1}  \hat\Omega_{\beta\beta} (X'X)^{-1}  
\end{aligned}
\]

This standard error estimator is an application of GLS that accounts for
the heteroskedasticity the model uses to identify the quantile
coefficients. We can apply the same principle to find the GLS standard
errors for the system of location-scale and quantile coefficients. To do
this, we define the following modified influence functions:

\[\begin{aligned}
\tilde \lambda_{1,i}&=\tilde \lambda_{2,i}=N (X'X)^{-1}  x_i ( x_i'\gamma) \\  
\tilde \lambda_{3,i}&=x_i' \gamma  \\
 \psi_{i,1} &= \varepsilon_i \\
 \psi_{i,2} &= \tilde \varepsilon_i -1 \\
 \psi_{i,3} &= \frac{1}{ x_i' \gamma }\frac{\tau-\mathbb{1}\big( q_\tau  \geq \varepsilon_i  \big) }{ f_{\varepsilon}(q_\tau)} - \frac{\varepsilon_i }{\bar x_i'\gamma} 
-  q_\tau \frac{(\tilde \varepsilon_i -1\big)}{\bar x_i'\gamma}
\end{aligned}
\]

Then, the GLS standard errors for the location-scale and quantile
coefficients can be estimated as

\[\hat{V}_{gls}
  \begin{pmatrix}
  \hat\beta \\
  \hat\gamma \\
  \hat q_\tau
  \end{pmatrix}
 = \frac{1}{N^2} 
 \begin{pmatrix}
 \hat\sigma_{11} \hat\Omega_{11} & \hat\sigma_{12} \hat\Omega_{12} & \hat\sigma_{13} \hat\Omega_{13} \\
 \hat\sigma_{12} \hat\Omega_{12} & \hat\sigma_{22} \hat\Omega_{22} & \hat\sigma_{23} \hat\Omega_{23} \\
 \hat\sigma_{13} \hat\Omega_{13} & \hat\sigma_{23} \hat\Omega_{23} & \hat\sigma_{33} \hat\Omega_{33}
  \end{pmatrix}
\]

where

\[\begin{aligned}
\hat\Omega_{jk} &= \frac{1}{N} \sum_{i=1}^N \tilde \lambda_{i,j}(\theta) \tilde \lambda_{i,k}(\theta)' \\
\hat\sigma_{jk} &= \frac{1}{N} \sum_{i=1}^N \psi_{i,j} \psi_{i,k}
\end{aligned}
\]

This estimator of standard errors is equivalent to the one derived by
MSS using Theorem 3. Empirically, the simulation study shows that this
estimator may be very sensitive to predictions of the scale model. In
the simulation study, we show that when the sample small, and the
likelihood of predicting zero or negative values for the scale model is
high, the GLS standard errors may be unreasonably large.

\subsection{\texorpdfstring{Multiple Fixed Effects: Expanding on
\citet{mss2019}}{Multiple Fixed Effects: Expanding on @mss2019}}\label{multiple-fixed-effects-expanding-on-mss2019}

Using the setup described in the previous section, MSS proposes an
extension to the model proposed by \citet{he1997} that enables the
estimation of quantile regression models with panel data, allowing for
the inclusion of individual fixed effects. However, the methodology can
also be generalized to allow for the inclusion of multiple fixed
effects. This type of analysis can be useful when considering data such
as employer-employee linked data \citep{abowed2006} or teacher-student
linked data \citep{harrissass2011}. Or, in the most common case,
allowing to control for both individual and time fixed effects.

We return to the original model and now assume there are sets of
unobserved heterogeneity that are constant across observations, if they
belong to common groups. Without loss of generality, we can assume that
the data-generating process is as follows:

\[\begin{aligned}
  y_{i} &= x_{i}' \beta + \delta_{g1} + \delta_{g2} + \nu_i \\
  \nu_i &= \varepsilon_i \times (x_{i}' \gamma + \zeta_{g1} + \zeta_{g2})   
  \end{aligned}
\]

where we assume \(x_{i}\) vary across groups \(g_1\) and \(g_2\) (and
thus are not collinear) and that \(\delta's\) and \(\zeta's\) are the
location and scale effects associated with the groups fixed
effects.\footnote{We could just as well consider multiple sets of fixed
  effects.}

If the dimension of groups \(g_k\) is low, this model could be estimated
using a dummy inclusion approach following Section~\ref{sec-betas}, and
the standard errors obtained as discussed in Section~\ref{sec-se}.
However, if the dimensionality of \(g_k\) is high, the dummy inclusion
approach may not be computationally feasible. A more feasible approach
is to apply the FWL theorem and partial out the impact of the group
fixed effects on the control variables \(x_{i}\), and the outcome of
interest \(y_{i}\), and using a similar approach for the identification
of \(\sigma(x)\). In the case of unbalanced setups with multiple groups,
the estimation involves iterative processes for which various approaches
have been suggested and implemented (see for example,
\citet{correia_feasible_nodate}, \citet{gaure2013}, \citet{rios2015},
among others).

When applying the partialing-out approach, some modifications to the
approach described in Section~\ref{sec-betas} are needed.

First, for all dependent and independent variables in the model
(\(w=y,x\)), we partial out the group fixed effects and obtain the
centered-residualized variables:

\[\begin{aligned}
w_{i} &= \delta_{g1}^w + \delta_{g2}^w + u_{i}^w \\
w_{i}^{rc} &= E(w_{i}) + \hat{u}_{i}^w
\end{aligned}
\]

Afterward, we estimate the location model using the
centered-residualized variables:\footnote{Using centered-residualized
  variables allows us to include a constant in the model specification,
  which simplifies the derivation of the influence functions. However,
  as with other fixed effects models, the constant is not identified and
  thus should not be interpreted.}

\[y_{i}^{rc} = x_{i}^{rc'} \beta + \nu_{i}
\]

Because \(|\hat \nu_i|\) is the dependent variable for the scale model,
we apply the partialing out and recentering to this expression
(\(|\hat \nu_i|^{rc}\)), and use that to estimate the following model:

\[|\hat\nu_{i}|^{rc} = x_{i}^{rc'} \gamma + \omega_{i}\]

Finally, the standardized residuals \(\varepsilon_i\) can be obtained as
follows:

\[\hat{\varepsilon}_{i} = \frac{\nu_{i}}{|\hat\nu_{i}|- \hat \omega_{i}}\]

where \(|\hat\nu_{i}|- \hat \omega_{i}\) is the prediction for the
conditional standard deviation
\(\sigma(x_i)=x_{i}' \gamma + \zeta_{g1} + \zeta_{g2}\)

The \(\tau_{th}\) quantile of the error \(\varepsilon\) can be estimated
as usual and the variance-covariance matrices obtained in the same way
as before (see Section~\ref{sec-se}) by using \(x_{i}^{rc}\) instead of
\(x_{i}\) when estimating the influence functions for all estimated
coefficients.

\section{Simulation Evidence}\label{simulation-evidence}

To show the performance of the extended strategy, we implement a small
simulation study. We consider a simple model with a single explanatory
variable \(x\). In contrast with MSS, we consider a two-way fixed effect
structure. For this exercise, we consider the following data-generating
process:

\[y_i = \alpha_{1i} + \alpha_{2i} + x_i + (2+x_i + \alpha_{1i} + \alpha_{2i} ) \varepsilon_i
\]

where \(\alpha_{1i}\sim \chi^2(1)\), \(\alpha_{2i} \sim \chi^2(1)\), and
\(x_i = 0.5 * (\chi_i +0.5(\alpha_{1i}+\alpha_{2i}))\), with
\(\chi_i \sim \chi^2(1)\). We only consider the case when the error term
\(\varepsilon_i\) is assumed to follow a centered \(\chi^2\)
distribution.\footnote{Specifically we assume
  \(\varepsilon=\frac{r_i}{5}-1\), where \(r_i\) follows a chi-squared
  distribution such that \(r_i \sim \chi^2(5)\). Simulations under the
  assumption of normal errors are available upon request.} We assume
that there are 50 mutually exclusive groups for each set of fixed
effects. Observations are assigned to each subgroup randomly using a
uniform distribution between 1 and 50.

To assess the alternative Standard error estimators, we consider a
second data-generating process where the error term is correlated within
clusters. For this, we assume observations are assigned ramdomly to 100
mutually exclusive groups, which are independent from the fixed effect
groups. In this setup the data-generating process is:

\[\begin{aligned}
y_i &= \alpha_{1i} + \alpha_{2i} + x_i + (2+x_i + \alpha_{1i} + \alpha_{2i} )*\kappa_i \\
\kappa_i &= inv-\chi^2_5(r_{i})/5-1 \\
r_i &= \Phi((\sqrt{.25}*s_{i}+\sqrt{.75}*s_{g}))
\end{aligned}
\]

where \(s_{i}\sim N(0,1)\) and \(s_{g}\sim N(0,1)\), are errors that
vary across individuals \(i\) or across clusters \(g\). \(\Phi()\) is
the cumulative distribution function of a normal distribution, and
\(inv-\chi^2_5()\) is the inverse function for a Chi-2 distribution with
5 degrees of freedom. With this setup, we generate an error structure
with a strong intra-cluster correlation, but without affecting the model
specification assumption.\footnote{We do not consider the case of
  misspecification of the scale model because that change would not only
  have an impact on the standard errors, but also affect the bias of the
  estimated coefficients}

We consider sample sizes of 500, 1000, 2000, and 4000 observations,
which implies an average of 10, 20, 40, and 80 observations per group.
The model is estimated using the location-scale model with
heteroskedastic linear errors and we report the coefficients for the
25th and 75th quantiles. We run this excercise 5000 times.
Table~\ref{tbl-sim1} reports the bias, simulated standard error, and
mean squared error, using the first data generation structure only. We
also report the results obtained using an adaptation bias-corrected
estimator based on the split-panel jackknife estimator proposed by
\citet{dhaene2015}.\footnote{For the implementation, we first estimate
  the model using the full sample, then randomly assign each observation
  into one of two groups, and finally reestimate the quantile
  coefficients for each group. The bias-corrected estimator is then
  obtained as
  \(\hat\beta(\tau)_{jkc}=2*\hat\beta(\tau)_{full}-0.5*\big(\hat\beta(\tau)_{s1}+\hat\beta(\tau)_{s2}\big)\).}
These results are labeled JKC.

Similar to the findings in MSS, we find that while the estimator
presents a substantial bias when the sample is small (\(N=500\)), this
bias shrinks as the sample size increases. As MSS describes, the bias
seems to be proportional to the sample size, or more precisely to the
average number of observations per sub-group. Interestingly, the
bias-corrected estimator presents an almost 0 bias for the 25th
percentile, even when the samples are small. In contrast, when
considering the 75th percentile, the simple estimator shows smaller bias
than the Jacknife estimator.{[}\^{}88: This is also consistent with
simulations with a single fixed effect{]} In either case, despite the
bias reduction obtained using the JKC estimator, the standard errors are
larger than without correction. For the 25th quantile, the reduction in
bias is large enough to produce a smaller Mean Squared Error (MSE) than
the simple estimator. This is similar to the results of MSS.

\begin{table}[H]

\caption{\label{tbl-sim1}Bias, Simulated Standard error, and Mean
Squared Error}

\centering{

\begin{tabular}{l*{6}{c}cccccccccccc}
\hline
            &     \multicolumn{3}{c}{N = 500}   &     \multicolumn{3}{c}{N = 1000}            \\
            &  Mean Bias &          SE&        MSE&   Mean Bias&          SE&        MSE \\
\hline
q25         &            &            &            &            &            &            \\
mmqreg      &       0.169&       0.267&       0.099&       0.092&       0.172&       0.038\\
jkc         &       0.048&       0.318&       0.104&       0.014&       0.189&       0.036\\
\hline
q75         &            &            &            &            &            &            \\
mmqreg      &      -0.050&       0.446&       0.202&      -0.010&       0.310&       0.096\\
jkc         &       0.048&       0.546&       0.301&       0.018&       0.339&       0.115\\
\hline
            &     \multicolumn{3}{c}{N = 2000}   &     \multicolumn{3}{c}{N = 4000}            \\
            &  Mean Bias &          SE&        MSE&   Mean Bias&          SE&        MSE \\
\hline
q25         &            &            &            &            &            &            \\
mmqreg      &       0.050&       0.119&       0.017&       0.026&       0.084&       0.008\\
jkc         &       0.006&       0.126&       0.016&       0.003&       0.087&       0.008\\
\hline
q75         &            &            &            &            &            &            \\
mmqreg      &       0.001&       0.215&       0.046&       0.003&       0.151&       0.023\\
jkc         &       0.006&       0.222&       0.049&       0.002&       0.154&       0.024\\
\hline
\multicolumn{7}{p{.7\textwidth}}{ \footnotesize Note: mmqreg - The proposed estimator. 
JKC-Jacknife Bias Corrected Estimator. SE- Simulated Standard Error.
MSE - Mean Squared Error. Mean bias is the difference 
between the estimated coefficient and the analytical true value. }\\
\end{tabular}

}

\end{table}%

To evaluate the performance of the different standard errors, we present
bias, 95\% coverage of the biased corrected estimates, as well as the
simulated standard errors, average and median of the standard errors
obtained using the GLS, robust SE, and clustered standard SE.
Table~\ref{tbl-sim2} considers the d.g.p. without intra-cluster
correlation, while Table~\ref{tbl-sim3} considers the d.g.p. that
induces intra-cluster correlations. All simulations consider 5000
repetitions.

The bias magnitude in Table~\ref{tbl-sim2} and Table~\ref{tbl-sim3} is
comparable to those observed in Table~\ref{tbl-sim1}, with simulated
standard errors that show to be slighly larger when we allow for
intra-cluster correlations, just as expected. When assuming there is no
need to clustered standard errors (Table~\ref{tbl-sim2}), the coverage
associated to GLS standard errors is above 95\% when the samples are
small, but it approximates to 95\% as the sample size increases. On the
other and, the coverage rates associated to Robust Standard errors are
closer to 90\%, albeit increasing slighly for larger samples.

One of the main reason that the GLS-Standard errors achieve higher than
expected rates of coverage may be related to the fact that the Standard
error estimator is very sensitive to near zero predictions from the
scale model. As shown in Table~\ref{tbl-sim2}, average and median GLS-SE
are considerably larger than the simulated standard errors when the
sample are small. While robust standard errors are less sensitive to
this problem, producing more stable results, they tend to underestimate
the magnitude of the true standard errors in small samples. This
translates into the lower coverage rates.

\begin{table}[H]

\caption{\label{tbl-sim2}95\% Coverage Ratio, and Standard Error
Estimation: No Intra-cluster Correlation}

\centering{

\begin{tabular}[hbt!]{l*{10}{c}cccccccccccccccccccc}
\hline 
            &      \multicolumn{4}{c}{Q25}   &      \multicolumn{4}{c}{Q75}   &\\
    N        &  =500& =1000 & =2000     &    =4000 &  =500& =1000 & =2000     &   =4000 & \\
\hline
Bias     &       0.173&       0.096&       0.050&       0.026&      -0.040&      -0.008&       0.003&       0.003\\
Sim SE         &       0.268&       0.171&       0.118&       0.084&       0.443&       0.314&       0.219&       0.153\\
Mean GLS SE      &1.9e7&       0.830&       0.401&       0.084&7.3e7 &       1.876&       4.259&       0.156\\
Median GLS SE     &       0.495&       0.215&       0.123&       0.083&       1.037&       0.429&       0.225&       0.152\\
CR95\%        &       0.988&       0.980&       0.958&       0.948&       0.991&       0.977&       0.967&       0.952\\
Mean Rbst SE     &       0.224&       0.159&       0.112&       0.080&       0.353&       0.269&       0.199&       0.144\\
CR95\%       &       0.892&       0.928&       0.939&       0.932&       0.875&       0.904&       0.927&       0.936\\
\hline 
%Mean CLS SE    &       0.235&       0.162&       0.112&       0.079&       0.372&       0.274&       0.199&       0.143\\
%CR95\%       &       0.907&       0.930&       0.932&       0.930&       0.892&       0.905&       0.925&       0.932\\    
\end{tabular}

}

\end{table}%

When we consider the presence of Intra-Cluster correlation,
Table~\ref{tbl-sim3}, we still observe similar problems with the GLS-SE,
albeit with high coverage rates. While Robust and Clustered Standard
errors are in average smaller than the Simulated Standard errors,
coverage rates are above 90\%. Clustered standard errors perform the
best only when the sample size is large, with Robust standard error
producing low Standard errors estimates.

\begin{table}[H]

\caption{\label{tbl-sim3}95\% Coverage Ratio, and Standard Error
Estimation: With Intra-cluster Correlation}

\centering{

\begin{tabular}[hbt!]{l*{10}{c}cccccccccccccccccccc}
    \hline 
             &      \multicolumn{4}{c}{Q25}   &      \multicolumn{4}{c}{Q75}   &\\
             N        &  =500& =1000 & =2000     &    =4000 &  =500& =1000 & =2000     &   =4000 & \\
   \hline
    Bias    &       0.180&       0.091&       0.053&       0.027&      -0.022&      -0.008&       0.003&       0.003\\
    Sim SE         &       0.299&       0.200&       0.141&       0.100&       0.503&       0.352&       0.253&       0.182\\
    Mean GLS SE      &1.5e7&       1.053&       0.190&       0.093&6.0e7 &       1.332&       0.309&       0.167\\
    Median GLS SE      &       0.476&       0.215&       0.133&       0.092&       0.913&       0.384&       0.236&       0.165\\
    CR95\%      &       0.984&       0.963&       0.939&       0.928&       0.986&       0.963&       0.940&       0.925\\
    Mean Rbst SE    &       0.253&       0.179&       0.126&       0.089&       0.402&       0.303&       0.222&       0.160\\
    CR95\%       &       0.896&       0.916&       0.917&       0.915&       0.880&       0.905&       0.917&       0.906\\
    Mean CLS SE    &       0.252&       0.180&       0.130&       0.096&       0.397&       0.304&       0.228&       0.172\\
    CR95\       &       0.892&       0.915&       0.923&       0.935&       0.875&       0.902&       0.919&       0.928\\
    \hline 
\end{tabular}

}

\end{table}%

Overall, the simulation study shows that using GLS-SE estimator may be
appropriate when the sample is relatively large. When there is suspected
presence of intra-cluster correlation, the use of clustered standard
errors is recommended in larger samples. However, when the sample is
small, the use of robust standard errors may be appropriate if GLS
standard errors when there is the risk of near-zero predictions from the
scale model.

\section{Illustrative applications}\label{illustrative-applications}

\subsection{Determinants of government
surpluses}\label{determinants-of-government-surpluses}

In this section we replicate one of the exercises from MSS, allowing for
time and individual fixed effects as well as for different standard
errors estimations. We use data from \citet{persson_economic_2005}, to
estimate the relationship between surplus of government as share of GDP,
and a measure of quality of democracy (POLITY); log of real income per
capita (LYP); trade volume as share of GDP (TRADE); share of population
between 15 and 65 years of age (P1564); the share of the population 65
years and older (P65); one-year lag of the dependent variable (LSP); oil
prices in US dollars, differntiating between importer and exporter
countries (OILIM and OILEX); and the output gap (YGAP). In addition to
country fixed effects (as illustrated in MSS), we show results allowing
for time fixed effects. Table~\ref{tbl-tb1} and Table~\ref{tbl-tb2}
provide the results for the model with and without time fixed effects,
respectively. The tables showcase the location and scale coefficients,
as well as the quantile coefficients for the 25th, 50th and 75th
quantiles. We also report GLS standard errors, robust standard errors
(brackets) and clustered standard errors at the country level.

\begin{table}[h]

\caption{\label{tbl-tb1}Determinants of Government Surpluses: Individual
Fixed Effects}

\centering{

    \centering
    \begin{tabular}{lccccccccc}
    \toprule
              & polityt &  lyp   & trade & p1564 &  p65   &  lspl  & oil-im & oil-ex &  ygap  \\
    \midrule
    \multicolumn{8}{l}{\textbf{Location}} \\
    \ \ \ coeff     &  0.116  & -0.715 & 0.030 & 0.121 & 0.028  & 0.691  & -0.047 & -0.006 & 0.010  \\
    \ \ \ se-gls   &  0.046  & 0.540  & 0.008 & 0.033 & 0.070  & 0.035  & 0.008  & 0.022  & 0.028  \\
    \ \ \ se-r     &  0.047  & 0.597  & 0.008 & 0.031 & 0.070  & 0.037  & 0.007  & 0.017  & 0.021  \\
    \ \ \ se-cl    &  0.046  & 0.465  & 0.007 & 0.032 & 0.071  & 0.035  & 0.010  & 0.020  & 0.023  \\
    \hline
    \multicolumn{8}{l}{\textbf{Scale}} \\
    \ \ \ coeff     & -0.097  & -0.616 & 0.003 & 0.036 & 0.087  & -0.085 & 0.013  & 0.016  & -0.004 \\
    \ \ \ se-gls   &  0.032  & 0.371  & 0.005 & 0.023 & 0.048  & 0.024  & 0.006  & 0.015  & 0.019  \\
    \ \ \ se-r     &  0.031  & 0.398  & 0.005 & 0.020 & 0.049  & 0.025  & 0.005  & 0.010  & 0.015  \\
    \ \ \ se-cl    &  0.048  & 0.800  & 0.008 & 0.031 & 0.067  & 0.029  & 0.004  & 0.010  & 0.012  \\
    \hline
    \multicolumn{8}{l}{\textbf{Q25}} \\
    \ \ \ coeff     &  0.191  & -0.239 & 0.028 & 0.093 & -0.039 & 0.756  & -0.057 & -0.018 & 0.013  \\
    \ \ \ se-gls   &  0.059  & 0.684  & 0.010 & 0.042 & 0.088  & 0.045  & 0.010  & 0.027  & 0.035  \\
    \ \ \ se-r     &  0.056  & 0.656  & 0.008 & 0.036 & 0.086  & 0.040  & 0.010  & 0.020  & 0.025  \\
    \ \ \ se-cl    &  0.073  & 0.687  & 0.006 & 0.041 & 0.098  & 0.023  & 0.010  & 0.021  & 0.029  \\
    \hline
    \multicolumn{8}{l}{\textbf{Q50}} \\
    \ \ \ coeff     &  0.108  & -0.765 & 0.030 & 0.124 & 0.035  & 0.684  & -0.046 & -0.005 & 0.009  \\
    \ \ \ se-gls   &  0.046  & 0.535  & 0.007 & 0.033 & 0.069  & 0.035  & 0.008  & 0.022  & 0.027  \\
    \ \ \ se-r     &  0.046  & 0.593  & 0.008 & 0.031 & 0.069  & 0.036  & 0.007  & 0.017  & 0.021  \\
    \ \ \ se-cl    &  0.043  & 0.484  & 0.008 & 0.032 & 0.070  & 0.036  & 0.010  & 0.020  & 0.023  \\
    \hline
    \multicolumn{8}{l}{\textbf{Q75}} \\
    \ \ \ coeff     &  0.031  & -1.258 & 0.033 & 0.153 & 0.104  & 0.616  & -0.036 & 0.008  & 0.006  \\
    \ \ \ se-gls   &  0.048  & 0.551  & 0.008 & 0.034 & 0.071  & 0.036  & 0.008  & 0.022  & 0.028  \\
    \ \ \ se-r     &  0.049  & 0.696  & 0.009 & 0.034 & 0.075  & 0.043  & 0.007  & 0.018  & 0.023  \\
    \ \ \ se-cl    &  0.039  & 0.919  & 0.012 & 0.041 & 0.079  & 0.055  & 0.010  & 0.022  & 0.020  \\
    \bottomrule
    \end{tabular}

}

\end{table}%

As expected, Table~\ref{tbl-tb1} shows that the point estimates are
identical to those reported in \citet{mss2019} (Table 6), including
analytical standard errors (GLS). With our estimator, however, we are
also able to produce both robust and clustered standard errors for
location and scale coefficients. Except for a few cases, the robust and
clustered standard errors are larger than the GLS standard errors, which
may be an indication of misspecification of the model. The GLS standard
errors we report differ from those in MSS, because they use panel
standard errors, which are equivalent to our clustered standard errors,
instead of the analytical standard errors we derive.

Considering the estimated effects across quantiles, we observe few
differences in the reported GLS standard errors compared to the
analytical standard errors reported in MSS. Our clustered standard
errors, however, are closer to the bootstrap-based standard errors the
authors report.\footnote{The differences in the GLS standard errors may
  be due that in our derivation the influence function of the
  standardized \(\tau_{th}\) quantile (see Equation~\ref{eq-ifs}) does
  not have the same leading term as the one reported in MSS (see Theorem
  3, and the definition of \(W\)).}

In Table~\ref{tbl-tb2}, we report the results including both individual
and year fixed effects. Because oil prices only vary across years, the
variable is excluded from the model specification. Accounting for time
fixed effects does not change the general conclusions that can be drawn,
based on the results from Table~\ref{tbl-tb1}. The two largest
differences are that the log of income per capita has a positive effect
on government surpluses, but only for the 25th quantile, because at this
point the largest impact on the Scale component is felt. Similarly, we
observe that the income gap has an impact on government surplus that is
always negative and increasing across quantiles. In both instances, the
effects are not statistically significant.

May be worth noting that the GLS SE are almost twice as large as the
robust and clustered standard errors. This is consistent with the
results from the simulation study. Examining the predictions of the
scale model, we find that there are 9 observations with a negative
predicted scale, and 9 observations that could be considered outliers.
This may be the reason for the large GLS standard errors.

\begin{table}[h]

\caption{\label{tbl-tb2}The Determinants of Government Surpluses:
Individual and Time Fixed Effects}

\centering{

    \centering
    \begin{tabular}{lccccccc}
    \hline
     & polity & lyp & trade & prop1564 & prop65 & lspl & ygap \\
    \hline
    \multicolumn{8}{l}{\textbf{Location}} \\
    \ \ \ coeff & 0.126 & -0.418 & 0.028 & 0.108 & 0.042 & 0.693 & -0.014 \\
    \ \ \ se-gls & 0.087 & 1.157 & 0.015 & 0.072 & 0.136 & 0.066 & 0.053 \\
    \ \ \ se-r & 0.047 & 0.703 & 0.008 & 0.038 & 0.068 & 0.038 & 0.022 \\
    \ \ \ se-cl & 0.048 & 0.506 & 0.008 & 0.044 & 0.077 & 0.037 & 0.022 \\
    \hline
    \multicolumn{8}{l}{\textbf{Scale}} \\
    \ \ \ coeff & -0.095 & -1.255 & 0.005 & 0.033 & 0.040 & -0.081 & 0.008 \\
    \ \ \ se-gls & 0.081 & 1.073 & 0.014 & 0.067 & 0.126 & 0.061 & 0.049 \\
    \ \ \ se-r & 0.031 & 0.452 & 0.005 & 0.025 & 0.045 & 0.025 & 0.017 \\
    \ \ \ se-cl & 0.041 & 0.848 & 0.006 & 0.030 & 0.048 & 0.033 & 0.013 \\
    \hline
    \multicolumn{8}{l}{\textbf{Q25}} \\
    \ \ \ coeff & 0.201 & 0.576 & 0.025 & 0.082 & 0.010 & 0.757 & -0.020 \\
    \ \ \ se-gls & 0.154 & 2.070 & 0.024 & 0.118 & 0.219 & 0.121 & 0.085 \\
    \ \ \ se-r & 0.058 & 0.751 & 0.008 & 0.049 & 0.080 & 0.040 & 0.026 \\
    \ \ \ se-cl & 0.073 & 0.761 & 0.006 & 0.052 & 0.087 & 0.023 & 0.027 \\
    \hline
    \multicolumn{8}{l}{\textbf{Q50}} \\
    \ \ \ coeff & 0.119 & -0.512 & 0.029 & 0.111 & 0.045 & 0.687 & -0.013 \\
    \ \ \ se-gls & 0.091 & 1.230 & 0.014 & 0.070 & 0.130 & 0.072 & 0.051 \\
    \ \ \ se-r & 0.046 & 0.695 & 0.008 & 0.037 & 0.068 & 0.038 & 0.022 \\
    \ \ \ se-cl & 0.045 & 0.529 & 0.008 & 0.044 & 0.077 & 0.039 & 0.021 \\
    \hline
    \multicolumn{8}{l}{\textbf{Q75}} \\
    \ \ \ coeff & 0.041 & -1.555 & 0.033 & 0.138 & 0.078 & 0.619 & -0.007 \\
    \ \ \ se-gls & 0.067 & 0.898 & 0.011 & 0.053 & 0.098 & 0.052 & 0.038 \\
    \ \ \ se-r & 0.048 & 0.827 & 0.009 & 0.037 & 0.075 & 0.046 & 0.026 \\
    \ \ \ se-cl & 0.038 & 0.980 & 0.012 & 0.050 & 0.086 & 0.063 & 0.020 \\
    \hline
    \end{tabular}

}

\end{table}%

\subsection{The Effect of Criminal Activities on Labor
Incomes}\label{sec-crime}

In this section, we investigate the effects of criminal activities on
labor incomes using a quantile regression approach to discover whether
crime has different impacts on the distribution of interest. We use data
from the ENEMDU labor survey from Ecuador and combine it with
administrative records of crime reports by Ecuadorian Cantons.

The combined data set covers the period 2019-2022. For the sake of
brevity, we will only present here the results of quantile regressions
of the log of total labor income on the log of annual reported murders
(LMURDERS) and other covariates displayed in Table 1. However, our
findings are robust to alternative specifications of the model where the
dependent variable is either the log of per-capita household income or
the log of total labor income from main occupation, both of which are
displayed in Section~\ref{sec-app-crime}.

\begin{table}[htbp]

\caption{\label{tbl-ecuador}The Effect of Crime on Labor Income: Canton,
Activity Sector and Time Fixed Effects}

\centering{

    \centering
    \begin{tabular}{lccccc}
    \toprule
    VARIABLES & q10 & q25 & q50 & q75 & q90 \\ \hline
     &  &  &  &  &  \\
    lmurders & -0.031* & -0.022* & -0.014* & -0.007 & -0.002 \\
     & (0.016) & (0.012) & (0.008) & (0.006) & (0.006) \\
    age & 0.062*** & 0.058*** & 0.054*** & 0.050*** & 0.047*** \\
     & (0.002) & (0.002) & (0.002) & (0.002) & (0.002) \\
    age$^2$ & -0.001*** & -0.001*** & -0.001*** & -0.001*** & -0.001*** \\
     & (0.000) & (0.000) & (0.000) & (0.000) & (0.000) \\
    male & 0.526*** & 0.435*** & 0.353*** & 0.284*** & 0.225*** \\
     & (0.049) & (0.037) & (0.027) & (0.020) & (0.017) \\
    informal & -0.841*** & -0.723*** & -0.617*** & -0.528*** & -0.451*** \\
     & (0.040) & (0.028) & (0.019) & (0.014) & (0.014) \\
    urban & -0.104*** & -0.102*** & -0.101*** & -0.100*** & -0.099*** \\
     & (0.014) & (0.012) & (0.011) & (0.012) & (0.013) \\
    married & 0.085*** & 0.094*** & 0.102*** & 0.108*** & 0.114*** \\
     & (0.013) & (0.010) & (0.007) & (0.006) & (0.005) \\
    educ & 0.026*** & 0.035*** & 0.042*** & 0.048*** & 0.053*** \\
     & (0.002) & (0.002) & (0.003) & (0.003) & (0.004) \\
    tenure & 0.010*** & 0.008*** & 0.007*** & 0.006*** & 0.005*** \\
     & (0.001) & (0.001) & (0.001) & (0.001) & (0.001) \\
    Constant & 3.715*** & 4.130*** & 4.500*** & 4.813*** & 5.082*** \\
     & (0.058) & (0.046) & (0.043) & (0.046) & (0.052) \\
     &  &  &  &  &  \\
    \bottomrule
    Observations & 445,521 & 445,521 & 445,521 & 445,521 & 445,521 \\
    Canton FE & Yes & Yes & Yes & Yes & Yes \\
    Activity Sector FE & Yes & Yes & Yes & Yes & Yes \\
    Year FE & Yes & Yes & Yes & Yes & Yes \\ 
    \bottomrule
    \multicolumn{6}{c}{ Clustered standard errors at Canton Level in parentheses} \\
    \multicolumn{6}{c}{ *** p$<$0.01, ** p$<$0.05, * p$<$0.1} \\
    \end{tabular}

}

\end{table}%

The findings in Table~\ref{tbl-ecuador} support the idea that the effect
of murders diminishes as we move forward in the distribution from low to
high incomes: in the lowest decile, a 1\% increase in reported murders
is associated with a 3.08\% reduction in total labor income; instead,
for the top 10\% labor income earners, murders are seen to have little
effect on the dependent variable and, in addition, the coefficient is
not significant at conventional levels. Quantile regression estimates
are thus really useful to discover relationships between dependent and
explanatory variables that are not seen in standard (mean) regression
analysis. For this application, we have found a monotonic decreasing
effect of the log of reported crimes on total labor income.

While the capability of estimating distinct coefficients of explanatory
variables for each quantile is a feature present since
\citet{koenkerbasset1978}, the method of moments approach allows us to
include not one but multiple sets of fixed effects so that one can
account for unobserved heterogeneity at different levels. Regarding this
application, we have included in the estimation fixed effects for
Canton, Activity Sector of the income earner and Year as is shown in
Table~\ref{tbl-ecuador}. Given that LMURDERS is measured at the Canton
level, it may be important to control for unobserved factors affecting
both crime and income earnings in the Cantons of Ecuador.

Finally, our approach also allows us to obtain clustered standard errors
for estimated coefficients so that statistical inference can be
performed properly. Table~\ref{tbl-ecuador} presents clustered standard
errors at the Canton level, the rationale being the same as described
above for the inclusion of Canton fixed effects in the specification.
Had we not taken into account the correlated nature of our standard
errors, we would have computed misleading (and smaller) estimates. For
example, robust standard errors in this specification are, on average,
2.5 times smaller than clustered standard errors. It is important,
however, to remark that the effects of murders on labor income at the
lower half of the distribution are still statistically different from
zero when clustering at the Canton level.

\section{Conclusions}\label{conclusions}

In this paper, we have extended the methodology proposed by
\citet{mss2019} in order to estimate quantile regression models with
multiple sets of fixed effects as well as with alternative standard
errrors. This methodology will allow researchers to implement
more-comprehensive analyses of data sets characterized by complex
hierarchies and unobserved heterogeneity. This extension is particularly
valuable in contexts where group specific effects vary across the
conditional distribution of the outcome of interest.

Using a small simulation study, we show that our extended approach is as
effective in identifying the parameters of interest as that of
\citet{mss2019}, even in contexts with two sets of fixed effects.
Notably, the bias-corrected estimator based on the split-panel jackknife
estimator exhibits promising results, mitigating biases when samples are
small but increasing standard errors.

Furthermore, we have assessed the impact of intracluster correlation on
the performance of standard error estimations. Our findings emphasize
the importance of using appropriate standard error estimators to ensure
accurate inference. In particular, we find that GLS standard errors are
biased when when the scale model predictions are close to zero or
negative, and that Robust and clustered standard errors are more stable
in those scenarios. Clustered Standard errors perform the best in the
presence of intra-cluster correlation, but the advantage is only evident
when the sample size is large.

Finally, we have illustrated the application of our extended methodology
using data from \citet{persson_economic_2005}. Our results are
consistent with those reported in \citet{mss2019}, and we are able to
provide robust and clustered standard errors for the location and scale
coefficients. We find that the GLS standard errors are almost twice as
large as the robust and clustered standard errors, which is consistent
with the results from our simulation study. Nevertheless, there is no
drastic change in the original conclusions.

Our example using data to analyze the impact of Crime on labor income in
Ecuador also suggests the relevance of using clustered standard errors
in the presence of unobserved heterogeneity at different levels. We find
that the effect of murders on labor income diminishes as we move forward
in the distribution from low to high incomes, which is a result that
would not have been possible without the use of quantile regression.

\section*{References}\label{references}
\addcontentsline{toc}{section}{References}

\renewcommand{\bibsection}{}
\bibliography{bibliography.bib}

\newpage{}

\appendix

\section{Derivation of the influence
functions}\label{derivation-of-the-influence-functions}

\subsection{Model Identification}\label{model-identification}

The estimation of quantile regression via moments assumes that the DGP
is linear in parameters, with an heteroskedastic error term that is also
a linear function of parameters:

\[\begin{aligned}
y_i = x_i'\beta + \nu_i \\
\nu_i = \varepsilon_i \times x_i'\gamma
\end{aligned}
\]

where \(\varepsilon\) is an unobserved i.i.d. random variable that is
independent of \(x\) and such that \(x\gamma\) is larger than 0 for any
\(x\).

In this case, the \(\tau_{th}\) conditional quantile model can be
written as

\[Q_y(\tau|X) = x'(\beta +Q_\varepsilon(\tau)  \times \gamma)\]

This model is identified under the following conditions:

\[\begin{aligned}
  E[(y_i-x_i'\beta)x_i ]  &= E[h_{1,i}]=0 \\
  E[ (|y_i-x_i'\beta|-x_i' \gamma) x_i ] &=E[h_{2,i}]=0 \\
  E\left[  \mathbb{1}\left( Q_\varepsilon(\tau) x_i'\gamma +x_i'\beta\geq  y_i  \right) - \tau \right] 
   &=E[h_{3,i}]=0 
\end{aligned}
\]

For simplicity, the rest of the appendix uses \(q_\tau\) to represent
\(Q_\varepsilon(\tau)\).

\subsection{Estimation of the Variance-Covariance
Matrix}\label{estimation-of-the-variance-covariance-matrix}

In this model, to estimate the variance-covariance matrix the set of
coefficients \(\theta'=[\beta' \ \gamma' \ q_\tau]\), we need to obtain
the influence functions of all coefficients, which are defined as

\[\lambda_i = \bar G(\theta)^{-1}
\begin{bmatrix}
h_{1,i} \\
h_{2,i} \\
h_{3,i}
\end{bmatrix}
\]

where the Jacobian matrix \(\bar G(\theta)\) is defined as

\[\bar G(\theta) = \begin{bmatrix}
\bar G_{11} & \bar G_{12} & \bar G_{13} \\
\bar G_{21} & \bar G_{22} & \bar G_{23} \\
\bar G_{31} & \bar G_{32} & \bar G_{13} \\
\end{bmatrix}
\]

with

\[\bar G_{j,k} = - \frac 1 N \sum_{i=1}^N \frac{\partial h_{j,i}}{\partial \theta_k'} \ \forall j,k \in 1,2,3
\]

First Moment Condition: Location Model

\[h_{1,i}=x_i(y_i-x_i'\beta)\]

\[\begin{aligned}
\bar G_{1,1} &=- \frac{1}{N} \sum_{i=1}^N \frac{\partial h_{1,i}}{\partial \beta'} \\
             &=- \frac{1}{N} \sum_{i=1}^N (-x_i x_i') \\
             &= N^{-1} X'X
\end{aligned}
\]

\[
\bar G_{1,2} = \bar G_{1,3} = 0
\]

Second Moment Condition: Scale model

\[h_{2,i}=x_i(|y_i-x_i'\beta|-x_i'\gamma)\]

\[\begin{aligned}
\bar G_{2,1} &= -\frac{1}{N} \sum \frac{\partial h_{2,i}}{\partial \beta'} \\
             &=  \frac{1}{N} \sum x_i x_i' \frac{y_i-x_i'\beta}{|y_i-x_i'\beta|} \\
\frac{y_i-x_i'\beta}{|y_i-x_i'\beta|} &= sign(y_i-x_i'\beta) \\            
\end{aligned}
\]

Under the assumption \(\varepsilon_i \times x\gamma\), or in this case
\(y_i-x_i'\beta\), is uncorrelated with \(x\), we can simplify the
expression as

\[\begin{aligned}
\bar G_{2,1} &= N^{-1} \left(N^{-1}\sum sign(y_i-x_i'\beta)\right) \sum x_i x_i' \\
&= N^{-1} E[sign(y_i-x_i'\beta)] X'X
\end{aligned}
\]

\[\begin{aligned}
\bar G_{2,2} &= -\frac{1}{N} \sum \frac{\partial h_{2,i}}{\partial \gamma'} \\
             &=  \frac{1}{N} \sum x_i x_i' \\
             &= N^{-1} X'X
\end{aligned}
\]

\[\bar G_{2,3}=0\]

Third Moment Condition: Quantile of Standardized Residual

\[\begin{aligned}
h_{3,i} &= \mathbb{1}\left( q_\tau  x_i'\gamma +x_i'\beta - y_i \geq 0 \right) - \tau \text{ or}\\
h_{3,i} &= \mathbb{1}\left( q_\tau  \geq \frac{y_i- x_i'\beta}{x_i'\gamma}  \right) - \tau = \mathbb{1}\big( q_\tau \geq \varepsilon  \big) - \tau \\
\end{aligned}
\]

Because the indicator function \(\mathbb{1}()\) is not differentiable,
we borrow from the nonparametric literature to approximate this function
with a kernel function. Call \(k()\) a well behaved kernel function that
is symetric around 0, and \(K()\) its integral, with range between 0 and
1. With an arbirarily small bandwidth \(h\), we can use the function
\(K()\) to approximate the indicator function:

\[
\lim_{h\rightarrow 0} K\left(\frac{z}{h}\right) \approx \mathbb{1}(z\geq 0)
\]

Thus the function \(h_{3,i}\) can be approximated as

\[h_{3,i} \approx K\left( \frac{1}{h} \Big( q_\tau x_i'\gamma +x_i'\beta - y_i  \Big) \right) - \tau
\]

Now, we can obtain the Jacobian matrix \(\bar G_{3,1}\) as:

\[\begin{aligned}
\bar G_{3,1} &= -\frac{1}{N} \sum \frac{\partial h_{3,i}}{\partial \beta'} \\
             &= -N^{-1} \sum \frac{1}{h} k\left( \frac{1}{h} \Big( q_\tau x_i'\gamma +x_i'\beta - y_i  \Big) \right) x_i' \\ 
             &= -N^{-1} \sum \frac{1}{h} k\left( \frac{1}{h} \big( q_\tau x_i'\gamma - \nu_i \big)   \right) x_i'  \\
             &= -N^{-1} \sum \frac{1}{h} k\left( \frac{1}{h} ( q_\tau - \varepsilon_i ) x_i'\gamma  \right) x_i'  
\end{aligned}
\]

If we condition on \(X\), use the law of iterated expectations, and
assume that \(\varepsilon_i\) is homoskedastic, we can obtain:

\[\begin{aligned}
\bar G_{3,1} &= -N^{-1} \sum \frac{1}{h} k\left( \frac{1}{h} ( q_\tau - \varepsilon_i ) x_i'\gamma  \right) x_i'  \\
             &= -N^{-1} \sum \frac{1}{h} k\left( \frac{q_\tau - \varepsilon_i}{h}  \right)  \frac{x_i'}{x_i'\gamma}  \\
             &= -N^{-1} f_\varepsilon (q_\tau) \sum  \frac{x_i'}{x_i'\gamma}  
\end{aligned}
\]

Finally, this simplifies to: \[\begin{aligned}
\bar G_{3,1} &\simeq -  f_\varepsilon (q_\tau) \frac{\bar x_i'}{ \bar x_i' \gamma} 
\end{aligned}
\]

where we use the fact that asymptotically, the expression
\(\frac{1}{N}\sum \frac{a_i}{b_i}\) can be approximated using Taylor
expansions by \(\frac{\bar a}{\bar b}\). \footnote{This approximation
  will be useful when we consider the estimation of the influence
  functions.} Thus, we can rewrite the last term as

The Jacobian for the second matrix \(\bar G_{3,2}\) can be derived
similarly:

\[\begin{aligned}
\bar G_{3,2} &= -\frac{1}{N} \sum \frac{\partial h_{3,i}}{\partial \gamma'} \\
             &= -N^{-1} \sum \frac{1}{h} k\left( \frac{1}{h} ( q_\tau - \varepsilon_i ) x_i'\gamma  \right) q_\tau x_i'  \\
             &= -N^{-1} \sum \frac{1}{h} k\left( \frac{q_\tau - \varepsilon_i}{h}    \right) q_\tau \frac{x_i'}{x_i'\gamma}  \\
             &\simeq  -  f(q_\tau)  q_\tau \frac{\bar x'}{\bar x'\gamma}
\end{aligned}
\]

and the Jacobian for the third matrix \(\bar G_{3,3}\) is

\[\begin{aligned}
\bar G_{3,3} &= -\frac{1}{N} \sum \frac{\partial h_{3,i}}{\partial q_\tau} \\
             &= -N^{-1} \sum \frac{1}{h} k\left( \frac{1}{h} ( q_\tau - \varepsilon_i ) x_i'\gamma  \right) x_i'\gamma  \\
             &= -N^{-1} \sum \frac{1}{h} k\left( \frac{q_\tau - \varepsilon_i}{h}   \right) \frac{x_i'\gamma}{x_i'\gamma}  \\
             &\simeq  -  f(q_\tau)  
\end{aligned}
\]

\subsection{Influence Functions}\label{influence-functions}

Location Coefficients

\[\lambda_i(\beta) = \bar G^{-1}_{1,1} h_{1,i} = N (X'X)^{-1}(x_i (y_i-x_i'\beta)) = N (X'X)^{-1}(x_i \nu_i)
\]

which can also be written as a function of the standardized residuals:

\[\lambda_i(\beta) = \bar G^{-1}_{1,1} h_{1,i} = N (X'X)^{-1}(x_i (y_i-x_i'\beta)) = N (X'X)^{-1}(x_i ( x_i'\gamma \times \varepsilon))
\]

Scale Coefficients

\[\begin{aligned}
\lambda_i(\gamma)&=\bar G_{2,2}^{-1}\Big(h_{2,i}-\bar G_{2,1} \lambda_i(\beta)\Big) \\
&=N (X'X)^{-1} \Big(x_i(|\nu_i|-x_i' \gamma) 
- N^{-1} E[sign(\nu_i)] X'X \big[ N (X'X)^{-1}(x_i \nu_i) \big] \Big) \\
&=N (X'X)^{-1} \Big(x_i(|\nu_i|-x_i' \gamma) - E[sign(\nu_i)] (x_i \nu_i)  \Big) \\
&=N (X'X)^{-1} x_i \Big( |\nu_i| - E[sign(\nu_i)] \nu_i -x_i' \gamma \Big) 
\end{aligned}
\]

However,

\[\begin{aligned}
|\nu_i| &= \nu_i \times \mathbb{1}(\nu_i \geq 0) - \nu_i \times \mathbb{1}(\nu_i < 0) \\
|\nu_i| &= \nu_i \times \mathbb{1}(\nu_i \geq 0) - \nu_i \times [1-\mathbb{1}(\nu_i \geq 0)] \\
|\nu_i| &= 2 \nu_i \times \mathbb{1}(\nu_i \geq 0) - \nu_i  
\end{aligned}
\]

and

\[\begin{aligned}
E[sign(\nu_i)] &= E[\mathbb{1}(\nu_i \geq 0)] - E[\mathbb{1}(\nu_i < 0)] \\
E[sign(\nu_i)] &= E[\mathbb{1}(\nu_i \geq 0)] - E[(1-\mathbb{1}(\nu_i \geq 0))] \\
E[sign(\nu_i)] &= 2 E[\mathbb{1}(\nu_i \geq 0)] - 1
\end{aligned}
\]

Thus,

\[\begin{aligned}
\lambda_i(\gamma) &= N(X'X)^{-1} x_i   \Big( 2 \nu_i \times \mathbb{1}(\nu_i \geq 0) - \nu_i  - ( 2 E[\mathbb{1}(\nu_i \geq 0)] - 1) \nu_i -x_i' \gamma \Big) \\
  &= N(X'X)^{-1} x_i   \Big( 2 \nu_i \times \mathbb{1}(\nu_i \geq 0) -  2 E[\mathbb{1}(\nu_i \geq 0)] \nu_i -x_i' \gamma \Big) \\
  &= N(X'X)^{-1} x_i   \Big( 2 \nu_i \times \big[ \mathbb{1}(\nu_i \geq 0) -  E[\mathbb{1}(\nu_i \geq 0)] \big] -x_i'\gamma \Big) \\
  &= N(X'X)^{-1} x_i   \Big( \tilde \nu_i -x_i' 
  \gamma \Big)
\end{aligned}
\]

This last expression is the equivalent simplification used in
\citet{mss2019} and \citet{im2000}. If the scale function is strictly
positive, it also follows that
\(\mathbb{1}(\nu_i \geq 0)= \mathbb{1}(\varepsilon_i \geq 0)\). Thus, it
can be simplified as

\[\begin{aligned}
\lambda_i(\gamma) &= N(X'X)^{-1} x_i ( x_i' \gamma ) \times (\tilde \varepsilon_i -1\big)
\end{aligned}
\]

Quantile of Standardized Residual

\[\begin{aligned}
\lambda_i(q_\tau)&=\bar G_{3,3}^{-1}
\Big(
 h_{3,i}-\bar G_{3,1} \lambda_i(\beta)-\bar G_{3,2} \lambda_i(\gamma)
\Big) \\
&=-\frac{1}{f_{\varepsilon}(q_\tau)} \times \Bigg( \Big(\mathbb{1} ( q_\tau  \geq \varepsilon  ) - \tau \Big)  \\
&+ f_{\varepsilon} (q_\tau) \frac{\bar x_i'}{\bar x_i'\gamma} 
N (X'X)^{-1} x_i  ( x_i'\gamma \times \varepsilon) \\
&+ f_{\varepsilon}(q_\tau) q_\tau \frac{\bar x_i'}{\bar x_i'\gamma} N(X'X)^{-1} x_i   \big( \tilde \nu_i -x_i' 
  \gamma \big) 
\Bigg) \\
&=\frac{\tau-\mathbb{1}\big( q_\tau  \geq \varepsilon  \big) }{f_{\varepsilon}(q_\tau)}
- \frac{ x_i'\gamma \times \varepsilon_i }{\bar x_i'\gamma} 
-  q_\tau \frac{ \tilde \nu_i -x_i' 
  \gamma }{\bar x_i'\gamma} 
\end{aligned}
\]

\section{\texorpdfstring{Complementary analysis of
Section~\ref{sec-crime}}{Complementary analysis of Section~}}\label{sec-app-crime}

In this appendix, we present the quantile regressions of both (log)
per-capita household income and the (log) of labor income from main
occupation on the (log) of annual reported murders. While in the first
specification the coefficients of interest are not significant at
conventional levels because of clustered standard errors (see
Table~\ref{tbl-acrime1}), the same pattern of decreasing association
between murders and measurements of well-being as we move to the right
of the distribution is revealed using our proposed method.

However, in Table~\ref{tbl-acrime2} we get significant (at the 1\%
level) associations between reported murders and a slightly different
measurement of labor income that only takes into account the main
occupation of the individual. Compared to total labor income, the
coefficients displayed in Table~\ref{tbl-acrime2} are smaller in
magnitude: for the lowest decile of the distribution, we find a 1
percentage point difference in the coefficient of interest between the
two alternative specifications, where the largest effect corresponds to
total labor income. Nevertheless, and once again, the decreasing pattern
of reported murders across quantiles is documented for this
specification. What is even more interesting, compared to our
estimations in Table~\ref{tbl-ecuador}, is that crime is no longer
statistically meaningful from the median of the distribution onwards.

\begin{table}[H]

\caption{\label{tbl-acrime1}The Effect of Crime on Per Capita Household
Income: Canton and Time Fixed Effects}

\centering{

    \centering
    \begin{tabular}{lccccc}
    \toprule
    VARIABLES & q10 & q25 & q50 & q75 & q90 \\ \hline
     &  &  &  &  &  \\
    lmurders & -0.026 & -0.022 & -0.019 & -0.015 & -0.012 \\
     & (0.019) & (0.015) & (0.012) & (0.009) & (0.008) \\
    age & 0.005*** & 0.005*** & 0.005*** & 0.006*** & 0.006*** \\
     & (0.001) & (0.001) & (0.001) & (0.001) & (0.001) \\
    age$^2$ & 0.000 & 0.000 & 0.000** & 0.000** & 0.000** \\
     & (0.000) & (0.000) & (0.000) & (0.000) & (0.000) \\
    male & -0.014** & -0.010** & -0.006 & -0.003 & 0.000 \\
     & (0.006) & (0.005) & (0.004) & (0.004) & (0.004) \\
    informal & -0.428*** & -0.377*** & -0.326*** & -0.279*** & -0.237*** \\
     & (0.029) & (0.024) & (0.018) & (0.014) & (0.011) \\
    urban & 0.136*** & 0.128*** & 0.119*** & 0.112*** & 0.105*** \\
     & (0.033) & (0.030) & (0.028) & (0.026) & (0.025) \\
    Constant & 5.046*** & 5.377*** & 5.707*** & 6.014*** & 6.282*** \\
     & (0.079) & (0.068) & (0.061) & (0.058) & (0.059) \\
     &  &  &  &  &  \\ \hline
    Observations & 566,127 & 566,127 & 566,127 & 566,127 & 566,127 \\
    Canton FE & Yes & Yes & Yes & Yes & Yes \\
    Year FE & Yes & Yes & Yes & Yes & Yes \\ \hline
    \multicolumn{6}{c}{\footnotesize Clustered standard errors at Canton Level in parentheses} \\
    \multicolumn{6}{c}{\footnotesize *** p$<$0.01, ** p$<$0.05, * p$<$0.1} \\
    \multicolumn{6}{c}{\footnotesize Coefficients for marital status, education attainment and household size not shown.} \\
\end{tabular}

}

\end{table}%

\begin{table}[H]

\caption{\label{tbl-acrime2}The Effect of Crime on Labor Income in Main
Occupation: Canton, Activity Sector and Time FE}

\centering{

    \centering
    \begin{tabular}{lccccc}
    \toprule
    VARIABLES & q10 & q25 & q50 & q75 & q90 \\ \hline
     &  &  &  &  &  \\
    lmurders & -0.021*** & -0.014** & -0.007 & -0.002 & 0.004 \\
     & (0.008) & (0.006) & (0.005) & (0.006) & (0.007) \\
    age & 0.061*** & 0.053*** & 0.046*** & 0.040*** & 0.034*** \\
     & (0.003) & (0.002) & (0.002) & (0.001) & (0.001) \\
    age$^2$ & -0.001*** & -0.001*** & -0.000*** & -0.000*** & -0.000*** \\
     & (0.000) & (0.000) & (0.000) & (0.000) & (0.000) \\
    male & 0.158*** & 0.149*** & 0.141*** & 0.133*** & 0.126*** \\
     & (0.018) & (0.015) & (0.012) & (0.010) & (0.010) \\
    informal & -0.515*** & -0.460*** & -0.410*** & -0.363*** & -0.321*** \\
     & (0.018) & (0.015) & (0.013) & (0.012) & (0.012) \\
    urban & -0.091*** & -0.098*** & -0.105*** & -0.111*** & -0.116*** \\
     & (0.012) & (0.011) & (0.010) & (0.010) & (0.011) \\
    married & 0.099*** & 0.093*** & 0.087*** & 0.082*** & 0.077*** \\
     & (0.007) & (0.005) & (0.004) & (0.004) & (0.004) \\
    educ & 0.042*** & 0.046*** & 0.050*** & 0.053*** & 0.056*** \\
     & (0.001) & (0.002) & (0.003) & (0.004) & (0.005) \\
    tenure & 0.007*** & 0.006*** & 0.005*** & 0.005*** & 0.004*** \\
     & (0.001) & (0.001) & (0.001) & (0.001) & (0.001) \\
    Constant & 3.882*** & 4.279*** & 4.649*** & 4.987*** & 5.294*** \\
     & (0.064) & (0.045) & (0.047) & (0.057) & (0.068) \\
     &  &  &  &  &  \\ \hline
    Observations & 249,993 & 249,993 & 249,993 & 249,993 & 249,993 \\
    Canton FE & Yes & Yes & Yes & Yes & Yes \\
    Activity Sector FE & Yes & Yes & Yes & Yes & Yes \\
    Year FE & Yes & Yes & Yes & Yes & Yes \\ \hline
    \multicolumn{6}{c}{ Clustered standard errors at Canton Level in parentheses} \\
    \multicolumn{6}{c}{ *** p$<$0.01, ** p$<$0.05, * p$<$0.1} \\
    \end{tabular}

}

\end{table}%

\section{Implementation}\label{implementation}

The method described here can be implemented using any of the following
packages:

\begin{itemize}
\tightlist
\item
  \texttt{mmqreg} in Stata:
  \texttt{net\ install\ mmqreg,\ from(https://friosavila.github.io/stpackages)}
\item
  \texttt{mmqreg} in R: Available on GitHub
  (https://github.com/friosavila/mmqreg)
\end{itemize}

\end{document}